\section{Discussion}

\subsection{Validity Concerns}

\subsubsection{Internal Validity}

The validity of the proof of concept was affected by several factors. Firstly, the experiment was limited to just three tools, mv, cp, and a self-made internal too. This may have impacted the results of the evaluation and reduced the chance of incorrectly retrieving command options, as a real-life scenario would not have allowed the tool to have such specialisation, for realistically, a user in a company may be asked to utilize several hundreds of tools. Secondly, the tool relies on the commands to have a sufficiently maintained catalog and documentation, which may not be a realistic expectation for real life constraints. Lastly, a degree of non-determinism was discovered during the testing phase. 

\subsubsection{External Validity and Generalizability}

The results of the study were affected by several limitations. Firstly, only bi-encoder embedding models were considered during model selection. Cross-encoder models may provide more accurate results because they process the query and candidate description together, but they are also more computationally expensive~\cite{qin2026ricci}. The results therefore cannot be directly applied to other retrieval approaches. Secondly, only three prompt variations were tested for the Qwen model family. Since small changes in prompt wording can affect the performance of an \ac{LLM}, the selected prompt may not represent the best possible performance of the model. Lastly, the user study included only nine participants. Although their feedback provided useful first impressions of the system, the sample size was too small to support general conclusions about user acceptance or usability. The user study should therefore be understood as an initial evaluation of the proof of concept rather than a representative study of a larger user group.

\subsubsection{Construct Validity and Evaluation Limitations}

The evaluation method also introduced several limitations. Firstly, the benchmark used a strict binary definition of success. A result was counted as incorrect whenever the expected tool was not retrieved. However, another tool or a different combination of options may still have completed the requested task. The benchmark may therefore have classified some usable results as failures. Secondly, all retrieval engines contributed to the final result using fixed weights. This does not account for the possibility that one engine may perform better than another for certain types of queries. Testing and calibrating different weights could improve the combined retrieval results. Finally, the benchmark mainly measured whether the predefined expected tools and options were retrieved. It did not fully measure whether the generated command would achieve the user’s intended result in practice. The reported metrics should therefore be interpreted as measures of retrieval performance rather than complete measures of real-world task success.

\subsection{Answers To The Research Questions}

\subsubsection{RQ1: Local-First and Privacy-Preserving System Design}

\textit{How can a local-first, privacy-preserving command option retrieval system be designed and implemented to run efficiently under strict local-hardware constraints?}

This question was addressed by designing and implementing a stateless, daemonless CLI tool using Rust. The system relies on a Hexagonal Architecture, which decouples core coordination, ranking, and validation rules from external adapters such as command-line entry points and persistent storage. To process natural language queries offline, the system integrates a dual-engine matching configuration. This setup combines a quantized semantic embedding model, specifically the eight-bit quantized variant of Google's three-hundred-million-parameter Gemma embedding model, with a BM25 lexical keyword search. Retrieval vector calculations are stored and compared using SQLite paired with the sqlite-vec extension. Syntactic safety is enforced by a Transition-Driven Command Assembly and Validation engine, which dynamically processes options, reorders command tokens, and filters out mutually exclusive flags. To ensure computational speed and zero external runtime dependencies on standard hardware, the binary is compiled targeting musl libc using Zig. This configuration embeds hardware-specific vector extensions, including AVX, AVX2, FMA, F16C, and BMI2, which allow the CPU to perform high-speed matrix multiplications and data unpacking. The resulting proof of concept runs completely offline, requires no background service daemons, and operates strictly within the target resource boundaries.

\subsubsection{RQ2: Retrieval Accuracy and Performance Efficiency}

\textit{How does the proposed retrieval system perform in terms of matching accuracy and runtime efficiency?}

The technical evaluation of the proof of concept indicates that the system achieves exceptional performance regarding computational efficiency, while its matching accuracy is moderate, making it suitable as a supportive learning aid rather than an autonomous execution engine.

From a performance perspective, the system operates well within the designated resource boundaries. It recorded an average query latency of 563.21~ms, with 99.18\% of all runs executing under the one-second limit. The peak memory consumption was highly stable and averaged 658.62~MB, comfortably staying below the 1024~MB constraint. These metrics confirm that the compiled binary and hardware instructions successfully minimize local computational overhead.

In terms of matching accuracy, the results are mixed. The classification model correctly routed queries to the target command-line utility in 80.30\% of cases ($N=489$). Within this correctly routed group, the command generation resolved successfully in 60.94\% of cases ($N=298$), while 39.06\% ($N=191$) failed validation due to option conflicts. For the 120 queries routed to the wrong tool, the validation engine successfully intercepted the routing error and blocked command generation in 53.33\% of cases ($N=64$). However, the system still confidently generated a command for the wrong tool in the remaining 46.67\% of these misrouted cases ($N=56$). Among the successfully resolved correct-tool commands, 24.16\% were perfect matches with 100\% option recall, 51.01\% were partial matches containing a subset of the required option flags, and 24.83\% returned zero correct options. Globally, this results in an incorrect command suggestion rate of 21.35\% across all queries, comprising 9.20\% where a wrong tool command was generated and 12.15\% where the correct tool was selected but with zero correct option flags. Thus, while the validation engine provides an effective safety filter, the rate of incorrect recommendations requires users to verify commands manually before execution.

\subsubsection{RQ3: Usability and Perceived Workflow Efficiency}

\textit{How do command-line interface (CLI) users of varying expertise levels evaluate the usability, helpfulness, and perceived workflow efficiency of the system?}

The user study, which included beginners, intermediate, and advanced command-line users, demonstrated that the tool significantly improves perceived workflow efficiency and reduces search effort. Across all expertise cohorts, participants valued the direct output of commands and the integration of placeholder explanations. A pronounced learning effect was observed during the study: the median completion time fell from 603 seconds for the first task to 214 seconds by the third task as participants became familiar with the interface. Usability on unfamiliar tools was also confirmed, as the median completion times for the custom, undocumented command-line utility ranged from 218 to 253 seconds, showing no systematic increase compared to familiar standard tools. Technology acceptance was high, with most participants agreeing that the tool is a valuable offline alternative in restricted environments. However, user confidence in the retrieved recommendations was less consistent, reflecting safety concerns about executing generated code in production environments. This highlight indicates that while the system accelerates workflows, integrating detailed explanations and transparent parameter descriptions is essential to foster user trust.
