\begin{appendix}

\section{Dataset Sample Categories}\label{app:dataset_sample_cats}
\begin{itemize}
    \item mv -iv /source/path /destination/path -- Moves all files from the source folder to the destination folder. The system checks before overwriting any matching names, asks for confirmation if matches are found and displays a live text confirmation for every single successful file transfer.
    \item mv -nv /source/path /destination/path -- Transfers items to the destination folder while strictly skipping any files that already exist there. It prints a list of the newly added items.
    \item mv -uv /source/path /destination/path -- Compares file modification times and transfers only new items or items modified since the last run. It ignores identical or newer destination files and provides a clear list of updates.
    \item cp -nR /source/path /destination/path -- Merges the contents of the source folder into the destination folder. It traverses all subdirectories to copy files but leaves existing files in the destination untouched, preventing any overwrites.
    \item cp -uri /source/path /destination/path -- Syncs the source to the destination by scanning all subfolders and copying only new or modified files. It still requests manual confirmation before overwriting any altered destination file.
    \item cp -sR /source/path /destination/path -- Recursively recreates a folder structure inside a project directory using symbolic links instead of duplicating actual data. It provides instant file access across directories with zero disk space cost.
    \item cp -fR /source/path /destination/path -- Forces a recursive copy of source files into the target directory. It automatically removes and replaces write-protected destination files without halting or asking for permission.
    \item mv --backup=numbered /source/path /destination/path -- Replaces the current file with the new file. Instead of destroying the original, it automatically renames it with a sequential number extension, ensuring a safety copy remains on disk.
    \item mv -f /source/path /destination/path -- Forces a complete transfer of files into the destination directory. It suppresses all confirmation prompts and overwrites existing files instantly.
    \item cp -av /source/path /destination/path -- Recursively clones the entire source folder to the destination folder. It preserves all original file permissions, ownership data, and modification timestamps while listing every copied file.
\end{itemize}

\section{Dataset Creation Prompt}\label{app:dataset_creation_prompt}

\begin{lstlisting}
You are an expert data generator for a Natural Language to Shell Command CLI tool. 

Your task is to generate exactly 10 realistic, diverse user queries based on a provided shell command and its technical description. These queries must perfectly mimic how a broad spectrum of users—ranging from total novices, intermediates, real human developers, system administrators, or tech-savvy users would naturally phrase their requests to an AI assistant.

CRITICAL: THE ALL-OR-NOTHING COMPLETENESS RULE
Every single generated query MUST capture the FULL intent of the command, including ALL side effects, flags, and behaviors mentioned in the description. You are strictly forbidden from leaving out details. If the command includes MULTIPLE flags, every query must find a way to express ALL requirements!

CRITICAL: ANTI-ANCHORING RULE
Do not mirror, copy, or mimic the sentence structures, specific words, or examples used in this prompt, it's examples or the provided description. Use entirely fresh phrasing, metaphors, and vocabulary to express the user's intent.

Here are the strict guidelines for the generated queries:

1. High Versatility: Vary the sentence structures. Mix imperative commands ("Move these..."), direct questions ("How can I..."), conversational requests ("I need to..."), and fragmented, task-oriented thoughts ("safely transfer files with output").
2. Synonyms & Vocabulary: Do not just copy words from the description. Use diverse phrasing (e.g., instead of "destination folder", use "target directory", "new folder", "another path").
3. Human Imperfections: Real users are sometimes terse, sometimes overly polite, and often informal. Ensure the tone spans from brief/direct to fully articulated sentences. 
4. Maximum Length: Every query must be a maximum of 1 to 2 short sentences.
5. Contextual Accuracy: The queries must strictly imply the technical nuances of the specific flags used in the command (e.g., if the command asks for confirmation or verbosity, the user query should subtly or explicitly request that safety/feedback).
6. Novice users might phrase the queries like someone who doesn't know command-line jargon. They use casual, everyday language (e.g., "send to", "dump into", "tell me what happened", "don't break anything"). They describe the *result* they want, not the technical process.
7. Intermediate users might Phrase the queries like someone who knows basic concepts but forgets exact flags. They use basic tech terms (e.g., "directory", "overwrite", "verbose", "prompt me").
8. Advanced users might Phrase the queries like a busy professional. Short, direct, imperative, or highly specific about the safety/output constraints of the command.

OUTPUT FORMAT:

You must return ONLY a JSON object. Do not include any markdown formatting wrappers (like ```json), no explanations, and no extra text. 

The JSON must follow this exact schema:

{
  "command": "The exact input command provided to you",
  "queries": [
    "Query 1",
    "Query 2",
    ...
    "Query 10"
  ]
} 

---
TARGET SCENARIO:
Command: [NAME]
Description: [DESCRIPTION]
\end{lstlisting}



\section{User Pre-Evaluation Questionnaire}\label{app:pre-evaluation_questions}

\begin{enumerate}
    \item \textbf{What command do you use to display the absolute path of the directory you are currently in?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{cd}
        \item[\textbf{B)}] \texttt{dir}
        \item[\textbf{C)}] \texttt{pwd}
        \item[\textbf{D)}] \texttt{path}
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{You are in \texttt{/home/user/projects/data}. You type \texttt{cd ..} and press Enter. Where are you now?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{/home/user/projects}
        \item[\textbf{B)}] \texttt{/home/user}
        \item[\textbf{C)}] \texttt{/home/user/projects/data/..}
        \item[\textbf{D)}] An error will occur.
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{If a program is currently running in your terminal and you want to forcefully stop or terminate it, what keyboard shortcut do you press?}
    \begin{itemize}
        \item[\textbf{A)}] Esc
        \item[\textbf{B)}] \texttt{Ctrl+X}
        \item[\textbf{C)}] \texttt{Ctrl+C}
        \item[\textbf{D)}] \texttt{Ctrl+Z}
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{What is the primary purpose of the \texttt{cat} command?}
    \begin{itemize}
        \item[\textbf{A)}] To search for specific words inside a file.
        \item[\textbf{B)}] To print the contents of a file directly to the terminal.
        \item[\textbf{C)}] To combine two directories together.
        \item[\textbf{D)}] To view the system hardware catalog.
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{Which command would you use to search for the specific word ``error'' inside a text file named \texttt{app.log}?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{find "error" app.log}
        \item[\textbf{B)}] \texttt{search app.log --text="error"}
        \item[\textbf{C)}] \texttt{grep "error" app.log}
        \item[\textbf{D)}] \texttt{cat app.log | match "error"}
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{How do you change the permissions of a script named \texttt{run.sh} to make it executable by the current user?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{chmod +x run.sh}
        \item[\textbf{B)}] \texttt{chown root run.sh}
        \item[\textbf{C)}] \texttt{make run.sh executable}
        \item[\textbf{D)}] \texttt{attrib +e run.sh}
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{What does the pipe operator (\textbar) do in a command sequence like \texttt{cat data.txt | grep "failed"}?}
    \begin{itemize}
        \item[\textbf{A)}] It runs the first command, waits for it to finish, then runs the second.
        \item[\textbf{B)}] It takes the standard output (stdout) of the first command and passes it as the standard input (stdin) to the second.
        \item[\textbf{C)}] It writes the output of both commands into a shared temporary file.
        \item[\textbf{D)}] It runs both commands simultaneously in the background.
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{You want to continuously watch the bottom of a log file as new lines are written to it in real-time. What command achieves this?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{watch app.log}
        \item[\textbf{B)}] \texttt{less --stream app.log}
        \item[\textbf{C)}] \texttt{tail -f app.log}
        \item[\textbf{D)}] \texttt{cat --live app.log}
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{Using the \texttt{find} command, how do you locate all files ending in \texttt{.tmp} and immediately delete them within that exact same command?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{find . -name "*.tmp" | rm}
        \item[\textbf{B)}] \texttt{find . -name "*.tmp" -delete}
        \item[\textbf{C)}] \texttt{find . -type "*.tmp" --remove}
        \item[\textbf{D)}] \texttt{rm \$(find "*.tmp")}
        \item[\textbf{E)}] I don't know
    \end{itemize}

    \item \textbf{You have a comma-separated file named \texttt{users.csv}. How would you extract only the second column (e.g., the usernames) and print it to the terminal?}
    \begin{itemize}
        \item[\textbf{A)}] \texttt{cut -d',' -f2 users.csv}
        \item[\textbf{B)}] \texttt{grep --column=2 users.csv}
        \item[\textbf{C)}] \texttt{column -extract 2 users.csv}
        \item[\textbf{D)}] \texttt{awk '\{print column 2\}' users.csv}
        \item[\textbf{E)}] I don't know
    \end{itemize}
\end{enumerate}

\section{User Evaluation Tasks}\label{app:Evaluation_tasks}

\textbf{Task 1: Create duplicate}

\textbf{Scenario}: In your current location there is a folder called data (/tmp/data). Your task is to create an exact copy of the folder and all its contents, in the misc directory (/tmp/misc). To use the tool: command-finder query "your query".

\textbf{Verification Command}: verify.sh 1

\textbf{Solution}: cp -r /tmp/data /tmp/misc NOTE: Other command options may have also been correct in this scenario.


\textbf{Task 2: Backup}  

\textbf{Scenario}: You have important data that you need to create a safety copy of, in a file called benchmark (/tmp/results/benchmark). However there is already some data in the backup directory and you want to make sure the copy in (/tmp/backup) is numbered! To use the tool: command-finder query "your query".  

\textbf{Verification Command}: verify.sh 2  

\textbf{Solution}: cp --backup numbered /tmp/results/benchmark /tmp/backup NOTE: Other command options may have also been correct in this scenario

\textbf{Task 3: Move to misc directory}  

\textbf{Scenario}: There is a file called numbers.txt in your current directory (/tmp/numbers.txt). Your job is to get this file into the misc directory (/tmp/misc/numbers.txt), updating any existing numbers.txt files in this directory. To use the tool: command-finder query "your query".  

\textbf{Verification Command}: verify.sh 3  

\textbf{Solution}: mv --update=all /tmp/numbers.txt /tmp/misc NOTE: Other command options may have also been correct in this scenario.


\textbf{Task 4: Environment Initialization}  

\textbf{Scenario}: Your goal is to prepare your current folder for a production environment. Use the correct tool to create a new, secure production setup file in this directory. Ensure the file is created with the correct secure and production settings! To use the tool: command-finder query "your query".  

\textbf{Verification Command}: verify.sh 4  

\textbf{Solution}: ops-sync --genesis -m:secure --env prod  


\textbf{Task 5: Legacy Payload Injection}

\textbf{Scenario}: There is an existing file in your directory named data.txt that needs to be modified. You need to use the tool to add a legacy version marker to the very end of this file. Make sure you also trigger the tool's header rotation feature during this process so the file updates correctly without any errors or lost data. To use the tool: command-finder query "your query".  

\textbf{Verification Command}: verify.sh 5  

\textbf{Solution}: ops-sync --append data.txt --stream legacy-v2 --force-rotate


\textbf{Task 6: Data Sanitization}

\textbf{Scenario}: You need to clean up sensitive data inside the existing data.txt file by removing any lines that contain the word "PASSWORD". Use the tool to perform this cleanup directly inside the file, and make sure to point the tool's required export pipeline flag to a new log file named clean.log so the system can record that the cleanup took place. To use the tool: command-finder query "your query".  

\textbf{Verification Command}: verify.sh 6  

\textbf{Solution}: ops-sync -s data.txt -x clean.log --scrub "PASSWORD" 

\section{User Evaluation Questionnaire}\label{app:User evaluation questionnaire}

\begin{enumerate}
    \item The tool was easy to understand during the tasks.

    \item I felt confident using the information provided by the tool.

    \item Using the tool reduced the effort needed to understand command options.

    \item I felt frustrated while using the tool.

    \item Overall, I was satisfied with the experience of using the tool.

    \item In a restricted environment where cloud-based AI cannot be used, this local tool would be a highly effective alternative for complex command options.

    \item Based on how well the tool retrieved options for standard commands during the test, I am confident it would successfully help me navigate a niche or internally developed CLI tool.

    \item Compared to the alternatives in a restricted environment, such as reading internal wikis, checking man pages, or asking a colleague, using this tool would significantly speed up my workflow.

    \item If my daily job required interacting with undocumented or proprietary command-line applications, I would actively want this local tool installed on my machine.

    \item What was the most helpful aspect of the tool during the tasks?

    \item What was the most confusing, frustrating, or unreliable aspect of the tool?

    \item In what situation would you prefer this tool over man pages, internal documentation, or asking a colleague? In what situation would you not use it?
\end{enumerate}


\section{Interface specification}\label{app:interfaces}
** VERIFY ALL REQUIREMENT REFERNCES IN THIS CHAPTER!!! Maybe remove the architecture stuff under the tables!**


\begin{table}[htbp]
    \centering
    \caption{Interface Specification: UserCommandPort (Inbound / Driving Port)}
    \label{tab:user_command_port}
    \renewcommand{\arraystretch}{1.4} % Adds breathing room
    \begin{tabularx}{\textwidth}{|p{3.2cm}|X|l|X|}
        \hline
        \rowcolor{gray!10} % Optional: Adds a light gray background to the header if you use \usepackage{colortbl}
        \textbf{Method} & \textbf{Parameters} & \textbf{Return} & \textbf{Exceptions} \\
        \hline
        \texttt{resolveQuery} & \texttt{rawQuery: String} & \texttt{String} & \texttt{Disambiguation\-Required\-Exception} \\
        \hline
        \texttt{update\-Configuration} & \texttt{config: EndUser\-Configuration} & \texttt{Boolean} & \texttt{Invalid\-Configuration\-Exception}, \newline \texttt{Storage\-Connection\-Exception} \\
        \hline
        \texttt{read\-Configuration} & \texttt{None} & \texttt{EndUser\-Configuration} & \texttt{\texttt{Storage\-Connection\-Exception}} \\
        \hline
        \multicolumn{4}{|p{\dimexpr\textwidth-2\tabcolsep-2\arrayrulewidth\relax}|}{
            \small \textbf{Architectural Purpose:} The primary entry point for translating a user's natural language request into an executable command. If similarity scores fall below the minimum threshold, or if the resulting command violates catalog exclusivity rules, it throws an exception to trigger the interactive clarification loop (FR-1, FR-6, FR-7).
        } \\
        \hline
    \end{tabularx}
\end{table}
\FloatBarrier


\begin{table}[htbp]
    \centering
    \caption{Interface Specification: CatalogIngestionPort (Inbound / Driving Port)}
    \label{tab:catalog_ingestion_port}
    \renewcommand{\arraystretch}{1.4}
    \begin{tabularx}{\textwidth}{|p{3.2cm}|X|l|X|}
        \hline
        \rowcolor{gray!10}
        \textbf{Method} & \textbf{Parameters} & \textbf{Return} & \textbf{Exceptions} \\
        \hline
        \texttt{ingestCatalog} & \texttt{catalog: ToolCatalog}, \newline \texttt{authKey: String} & \texttt{Boolean} & \texttt{Authentication\-Exception}, \newline \texttt{Invalid\-Catalog\-Exception}, \newline \texttt{Storage\-Connection\-Exception} \\
        \hline
        \texttt{updateCatalog} & \texttt{catalog: ToolCatalog}, \newline \texttt{authKey: String} & \texttt{Boolean} & \texttt{Authentication\-Exception}, \newline \texttt{Catalog\-NotFound\-Exception}, \newline \texttt{Invalid\-Catalog\-Exception}, \newline \texttt{Storage\-Connection\-Exception} \\
        \hline
        \texttt{deleteCatalog} & \texttt{toolName: String}, \newline \texttt{authKey: String} & \texttt{Boolean} & \texttt{Authentication\-Exception}, \newline \texttt{Catalog\-NotFound\-Exception}, \newline \texttt{Storage\-Connection\-Exception} \\
        \hline
        \multicolumn{4}{|p{\dimexpr\textwidth-2\tabcolsep-2\arrayrulewidth\relax}|}{
            \small \textbf{Architectural Purpose:} Defines the administrative contract for managing the lifecycle of tool catalogs. The external CLI adapter handles JSON deserialization, while this port ensures that all catalog mutations are strictly authenticated and semantically valid before the core delegates them to the persistence layer (FR-3, FR-4, FR-5).
        } \\
        \hline
    \end{tabularx}
\end{table}
\FloatBarrier


\renewcommand{\arraystretch}{1.4}
% The column definition now automatically permits breaks anywhere inside \texttt strings
\begin{longtable}{|>{\raggedright\arraybackslash\hspace{0pt}\emergencystretch=1em\hyphenpenalty=1}p{3.2cm}|>{\raggedright\arraybackslash}p{3.5cm}|l|>{\raggedright\arraybackslash}p{4cm}|}
    
    % --- FIRST PAGE HEADER ---
    \caption{Interface Specification: StoragePort (Outbound / Driven Port)} \label{tab:storage_port} \\
    \hline
    \rowcolor{gray!10}
    \textbf{Method} & \textbf{Parameters} & \textbf{Return} & \textbf{Exceptions} \\
    \hline
    \endfirsthead

    % --- SUBSEQUENT PAGES HEADER ---
    \hline
    \rowcolor{gray!10}
    \multicolumn{4}{|l|}{\small\textit{Table \ref{tab:storage_port} continued from previous page}} \\
    \hline
    \rowcolor{gray!10}
    \textbf{Method} & \textbf{Parameters} & \textbf{Return} & \textbf{Exceptions} \\
    \hline
    \endhead

    % --- FOOTER FOR PAGES THAT SPLIT ---
    \hline
    \multicolumn{4}{|r|}{\small\textit{Continued on next page...}} \\
    \hline
    \endfoot
    
    % --- FINAL FOOTER ---
    \hline
    \endlastfoot

    % --- TABLE DATA ---
    \multicolumn{4}{|l|}{\textbf{Catalog Management}} \\
    \hline
    \texttt{save\-Catalog} & \texttt{catalog: ToolCatalog} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception}, \newline \texttt{Duplicate\-Catalog\-Exception} \\
    \hline
    \texttt{update\-Catalog} & \texttt{catalog: ToolCatalog} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception}, \newline \texttt{Catalog\-NotFound\-Exception} \\
    \hline
    \texttt{delete\-Catalog} & \texttt{toolName: String} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception}, \newline \texttt{Catalog\-NotFound\-Exception} \\
    \hline
    \texttt{fetch\-Catalog} & \texttt{toolName: String} & \texttt{ToolCatalog} & \texttt{Storage\-Connection\-Exception}, \newline \texttt{Catalog\-NotFound\-Exception} \\
    \hline
    \texttt{fetch\-All\-Catalogs} & \texttt{None} & \texttt{List<ToolCatalog>} & \texttt{Storage\-Connection\-Exception} \\
    \hline
    
    \multicolumn{4}{|l|}{\textbf{Authentication / Maintainer Data}} \\
    \hline
    \texttt{save\-Maintainer} & \texttt{maintainer: Catalog\-Maintainer} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception} \\
    \hline
    \texttt{update\-Maintainer} & \texttt{maintainer: Catalog\-Maintainer} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception, \newline \texttt{Maintainer\-NotFound\-Exception}} \\
    \hline
    \texttt{fetch\-Maintainer} & \texttt{maintainerId: String} & \texttt{Catalog\-Maintainer} & \texttt{Storage\-Connection\-Exception}, \newline \texttt{Maintainer\-NotFound\-Exception} \\
    \hline
    \texttt{delete\-Maintainer} & \texttt{maintainerId: String} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception}, \newline \texttt{Maintainer\-NotFound\-Exception}, \newline \texttt{Maintainer\-NotFound\-Exception} \\
    \hline
    
    \multicolumn{4}{|l|}{\textbf{User Configuration}} \\
    \hline
    \texttt{load\-Configuration} & \texttt{None} & \texttt{EndUser\-Config} & \texttt{Storage\-Connection\-Exception} \\
    \hline
    \texttt{save\-Configuration} & \texttt{config: EndUser\-Config} & \texttt{Boolean} & \texttt{Storage\-Connection\-Exception} \\
    \hline
    
    \multicolumn{4}{|p{\dimexpr\textwidth-2\tabcolsep-2\arrayrulewidth\relax}|}{
        \small \textbf{Architectural Purpose:} Defines the absolute contract for persisting and retrieving domain entities. Note that \texttt{fetchCatalog} inherently retrieves all associated \texttt{CommandDefinitions} and \texttt{ExclusivityRules} required by the Syntactical Validator. By strictly relying on this interface, the Application Core remains completely agnostic to the underlying database technology (e.g., SQLite).
    } \\
\hline
\end{longtable}
\FloatBarrier

\begin{table}[htbp]
    \centering
    \caption{Interface Specification: MatchingStrategyPort (Outbound / Driven Port)}
    \label{tab:matching_strategy_port}
    \renewcommand{\arraystretch}{1.4}
    \begin{tabularx}{\textwidth}{|p{3.2cm}|X|X|X|}
        \hline
        \rowcolor{gray!10}
        \textbf{Method} & \textbf{Parameters} & \textbf{Return} & \textbf{Exceptions} \\
        \hline
        \texttt{find\-Tools} & \texttt{query: UserQuery}, \newline & \texttt{List\-<List\-<Scored\-Tool>>} & \texttt{Engine\-Execution\-Exception}\\
        \hline
        \texttt{find\-Options} & \texttt{query: UserQuery, toolName: String}, \newline & \texttt{List\-<List\-<Scored\-Option>>} & \texttt{Engine\-Execution\-Exception}\\
        \hline
        \texttt{load\-Engine} & \texttt{None} & \texttt{Boolean} & \texttt{Initialization\-Exception} \\
        \hline
        \texttt{create\-Optimized\-Catalog} & \texttt{catalog: toolCatalog} & \texttt{None} & \texttt{Initialization\-Exception}, \texttt{Storage\-Connection\-Exception} \\
        \hline
        \texttt{update\-Optimized\-Catalog} & \texttt{catalog: toolCatalog} & \texttt{None} & \texttt{Initialization\-Exception}, \texttt{Storage\-Connection\-Exception} \\
        \hline
        \texttt{delete\-Optimized\-Catalog} & \texttt{toolName: String} & \texttt{None} & \texttt{Storage\-Connection\-Exception} \\
        \hline
        \texttt{toolWeight} & \texttt{None} & \texttt{Float} & \texttt{Initialization\-Exception} \\
        \hline
        \texttt{optionWeight} & \texttt{None} & \texttt{Float} & \texttt{Initialization\-Exception} \\
        \hline
        \multicolumn{4}{|p{\dimexpr\textwidth-2\tabcolsep-2\arrayrulewidth\relax}|}{
            \small \textbf{Architectural Purpose:} Defines the contract for evaluating query similarity against candidate commands. By exposing only this interface to the core, the system achieves true algorithmic decoupling (NFR-7). External adapters can implement this port using anything from simple lexical token-matching to advanced neural embeddings (NFR-8). Furthermore, the \texttt{preloadModels} method guarantees that heavy machine learning assets are loaded into RAM during the boot or JIT phase, thereby preserving sub-second query latency during actual execution (NFR-1, NFR-2). Additionally methods are provided to manage optimized catalog instances, which allows an engine to store the catalog in it's preprocessed engine specific form. This reduces computational overhead during execution, contributing to the sub-second query latency. The \texttt{toolWeight} and \texttt{optionWeight} methods allow to give more accurate engines a higher importance during the ranking of candidates.
        } \\
        \hline
    \end{tabularx}
\end{table}
\FloatBarrier

\section{Tool Catalogs}\label{app:tool_catalogs}

\subsection{cp-catalog.json}
\begin{lstlisting}
{
  "toolName": "cp",
  "description": "The copy tool is a filesystem utility designed to duplicate files, clone data, and replicate entire directory trees for backups or asset creation.",
  "userFriendlyDescription": "Copies files and folders from one location to another.",
  "keywords": "cp copy duplicate clone replica backup directory folder transfer recursive replicate archiving preserve symlink hardlink override overwrite data sync migrate source destination path",
  "version": "0.1",
  "options": [
    {
      "optionName": "-a",
      "description": "Parent: cp (filesystem management). Option: -a (--archive). Function: Archive deployment; exact replication via recursive symlink preservation and comprehensive attribute retention.",
      "userFriendlyDescription": "Copies files and folders while keeping all of their original properties, permissions, and shortcuts intact.",
      "keywords": "archive recursive recurse links metadata ownership timestamps permissions retention preserve-all backup-archive entire-directory-tree preserve-attributes attributes symbolic symlink keep-all-details keep-details entire-folder-contents"
    },
    {
      "optionName": "--attributes-only",
      "description": "passage: Parent: cp (filesystem management). Option: --attributes-only (). Function: Metadata replication; absolute restriction to file attributes with absolute omission of data payloads.",
      "userFriendlyDescription": "Creates empty files that copy only the settings and info of the original files without copying their contents.",
      "keywords": "attributes-only metadata-only attributes permissions ownership timestamps retention size skip-data omit-payload do-not-copy-content no-data-transfer metadata empty-files copy-only-attributes no-data"
    },
    {
      "optionName": "--backup",
      "description": "passage: Parent: cp (filesystem management). Option: --backup (=CONTROL). Function: Destination redundancy; generation of structured backups for existing destination targets.",
      "userFriendlyDescription": "Makes a backup copy of any file at the destination that would otherwise be overwritten.",
      "keywords": "backup control suffix versioning overwrite redundancy backups numbered-backup backup-method control-parameter replace duplicate exist collision conflict safety-copy save-old keep-old auto-backup versioned-backups"
    },
    {
      "optionName": "-b",
      "description": "Parent: cp (filesystem management). Option: -b (). Function: Argument-free redundancy; generation of standard destination backups without parameter configuration.",
      "userFriendlyDescription": "Makes a standard backup of destination files before overwriting them without requiring extra settings.",
      "keywords": "backup default argumentless redundancy standard-backup simple-backup parameterless-backup argumentless-backup replace duplicate exist collision conflict overwrite safety-copy save-old keep-old auto-backup default-backup"
    },
    {
      "optionName": "--copy-contents",
      "description": "passage: Parent: cp (filesystem management). Option: --copy-contents (). Function: Special file payload replication; recursive processing of character, block, or pipe device data.",
      "userFriendlyDescription": "Copies the actual contents of special system files instead of just copying the files themselves.",
      "keywords": "copy-contents special-files fifo block character devices nodes pipes devices-contents special-device-nodes recursive copy-special-device-payloads"
    },
    {
      "optionName": "-d",
      "description": "passage: Parent: cp (filesystem management). Option: -d (). Function: Link structural replication; absolute symbolic link retention and preservation across targets.",
      "userFriendlyDescription": "Copies shortcuts as shortcuts rather than copying the files they point to.",
      "keywords": "no-dereference symbolic symlink link shortcut preserve-symlinks link-replication keep-symbolic-links copy-symlinks-as-links resolution dereference copy-shortcut-itself keep-shortcuts"
    },
    {
      "optionName": "--debug",
      "description": "passage: Parent: cp (filesystem management). Option: --debug (). Function: Execution introspection; verbose diagnostics detailing copying mechanisms with implicit verbosity activation.",
      "userFriendlyDescription": "Shows detailed technical information to help troubleshoot how files are copied.",
      "keywords": "debug verbose explain diagnostics logging traces execution-introspection troubleshooting diagnose-copy diagnostics-telemetry"
    },
    {
      "optionName": "-f",
      "description": "passage: Parent: cp (filesystem management). Option: -f (--force). Function: Overwrite enforcement; absolute removal of unopenable destination files ahead of creation attempts.",
      "userFriendlyDescription": "Forces the copy to succeed by removing and replacing existing destination files that cannot be opened.",
      "keywords": "force overwrite silent automatic override bypass skip suppress ignore permissions protection clobber unconditional non-interactive automatically silently unopenable delete-unwritable remove deletion"
    },
    {
      "optionName": "-i",
      "description": "passage: Parent: cp (filesystem management). Option: -i (--interactive). Function: Confirmation intervention; mandatory user prompting prior to destination data overwrites.",
      "userFriendlyDescription": "Asks you for confirmation before overwriting any existing files.",
      "keywords": "interactive prompt confirmation ask confirm check verify warning safety user validation query question overwrite replacement input prompt-overwrite confirm-overwrite verify-overwrite interactive-prompt user-confirmation validation-prompt confirm-replacement safety-warning"
    },
    {
      "optionName": "-H",
      "description": "passage: Parent: cp (filesystem management). Option: -H (). Function: Source command-line symlink resolution; physical traversal of explicit command-line symbolic parameters.",
      "userFriendlyDescription": "Follows and copies the actual files for shortcuts specified directly in the command.",
      "keywords": "follow-command-line symbolic symlink link shortcut dereference resolution traversal-links command-line-arguments-only follow-explicit-links command-line-symlinks resolve-command-line-parameters"
    },
    {
      "optionName": "-l",
      "description": "passage: Parent: cp (filesystem management). Option: -l (--link). Function: Hard link creation instead of file replication.",
      "userFriendlyDescription": "Creates links pointing to the original files on the hard drive instead of copying their data.",
      "keywords": "hardlink link-creation pointer inode-reference inode pointer reference create-hard-links-instead link copy-hardlink"
    },
    {
      "optionName": "-L",
      "description": "passage: Parent: cp (filesystem management). Option: -L (--dereference). Function: Mandatory source symbolic link resolution.",
      "userFriendlyDescription": "Always follows shortcuts and copies the actual files they point to.",
      "keywords": "dereference follow-symlinks symbolic symlink link shortcut resolution traversal follow-symbolic follow-all-links copy-link-contents follow-shortcuts follow-all-shortcuts copy-target-content actual-files-instead-of-shortcuts"
    },
    {
      "optionName": "-n",
      "description": "passage: Parent: cp (filesystem management). Option: -n (--no-clobber). Function: Existing file protection; override of updates or interactive prompts.",
      "userFriendlyDescription": "Prevents overwriting any files that already exist at the destination.",
      "keywords": "no-clobber do-not-overwrite overwrite-protection skip-existing safety ignore-collision prevent-clobbering replace duplicate exist collision conflict dont-delete skip-duplicates prevent-overwrite keep-old-files"
    },
    {
      "optionName": "-P",
      "description": "passage: Parent: cp (filesystem management). Option: -P (--no-dereference). Function: Symbolic link preservation; strict restriction to source link entities.",
      "userFriendlyDescription": "Copies shortcuts as shortcuts and never follows them to copy the original files.",
      "keywords": "no-dereference preserve-symlinks symbolic symlink link shortcut copy-symlinks skip-resolution never-dereference-links resolution dereference never-follow-symlinks dont-follow-shortcuts"
    },
    {
      "optionName": "-p",
      "description": "passage: Parent: cp (filesystem management). Option: -p. Function: Default preservation of mode, ownership, and timestamp metadata.",
      "userFriendlyDescription": "Preserves the original file owner, permissions, and timestamps when copying.",
      "keywords": "preserve-metadata preserve-mode preserve-ownership preserve-timestamps preserve-permissions attributes metadata maintain-attributes permissions ownership timestamps retention keep-details keep-permissions keep-owner keep-dates creation-time"
    },
    {
      "optionName": "--preserve",
      "description": "passage: Parent: cp (filesystem management). Option: --preserve[=ATTR_LIST]. Function: Selective retention of specified file attributes.",
      "userFriendlyDescription": "Allows you to choose specific file properties, like ownership or timestamps, to keep during the copy.",
      "keywords": "preserve-attributes preserve-acl preserve-security preserve-context selective-preserve attributes permissions ownership timestamps metadata retention custom-preservation selective-metadata keep-acl keep-context security-attributes"
    },
    {
      "optionName": "--no-preserve",
      "description": "passage: Parent: cp (filesystem management). Option: --no-preserve]. Function: Selective omission of specified file attributes.",
      "userFriendlyDescription": "Allows you to choose specific file properties to discard during the copy.",
      "keywords": "no-preserve discard drop strip attributes permissions ownership timestamps metadata retention omit-attributes discard-metadata skip-permissions skip-metadata omit-permissions"
    },
    {
      "optionName": "--parents",
      "description": "passage: Parent: cp (filesystem management). Option: --parents. Function: Path structural replication under target directory.",
      "userFriendlyDescription": "Recreates the original folder structure inside the destination folder.",
      "keywords": "parents recreate-parents create-parent-directories path-structure directory-hierarchy directory-creation path-hierarchy recreate-subdirectories full-path create-folders"
    },
    {
      "optionName": "-R",
      "description": "passage: Parent: cp (filesystem management). Option: -R (-r, --recursive). Function: Recursive directory tree traversal and duplication.",
      "userFriendlyDescription": "Copies folders and all of their contents, including subfolders and files.",
      "keywords": "recursive recurse copy-directories copy-folder tree-traversal nested-folders entire-folder-contents copy-subfolders recursive-copy copy-directories-and-contents subdirectories everything-inside"
    },
    {
      "optionName": "--reflink",
      "description": "passage: Parent: cp (filesystem management). Option: --reflink[=WHEN]. Function: Copy-on-write clone behavior management.",
      "userFriendlyDescription": "Performs a quick, space-saving copy that only duplicates data when it is modified.",
      "keywords": "reflink clone cow copy-on-write btrfs zfs snapshot lightweight-copy copy-on-write-clone lightweight-clone"
    },
    {
      "optionName": "--remove-destination",
      "description": "passage: Parent: cp (filesystem management). Option: --remove-destination (). Function: Pre-emptive target file erasure prior to initialization.",
      "userFriendlyDescription": "Deletes any existing destination files before starting the copy.",
      "keywords": "remove-destination delete-existing unlink-before-open force-overwrite unlink-destination remove-before-copy replace duplicate exist collision conflict overwrite delete-before-write delete-first unlink-target"
    },
    {
      "optionName": "--sparse",
      "description": "passage: Parent: cp (filesystem management). Option: --sparse=[WHEN]. Function: Conditional sparse file allocation management.",
      "userFriendlyDescription": "Saves disk space when copying files by optimizing empty blocks of data.",
      "keywords": "sparse sparse-file hole-allocation space-saving efficiency sparse-creation dynamic-allocation dynamic-blocks"
    },
    {
      "optionName": "--strip-trailing-slashes",
      "description": "passage: Parent: cp (filesystem management). Option: --strip-trailing-slashes (). Function: Source argument trailing slash deletion.",
      "userFriendlyDescription": "Removes slash marks from the ends of folder names before copying them.",
      "keywords": "strip-trailing-slashes remove-trailing-slashes path-normalization trim-pathname strip-slashes remove-slash"
    },
    {
      "optionName": "-s",
      "description": "passage: Parent: cp (filesystem management). Option: -s (--symbolic-link). Function: Symbolic link generation substitution for data replication.",
      "userFriendlyDescription": "Creates shortcuts to the source files instead of copying their data.",
      "keywords": "symbolic-link symlink soft-link softlink shortcut make-shortcuts instead-copying create-symlinks-instead link resolution create-shortcuts softlink-generation"
    },
    {
      "optionName": "--suffix",
      "description": "passage: Parent: cp (filesystem management). Option: -S (--suffix). Function: Backup suffix identifier override.",
      "userFriendlyDescription": "Adds a custom ending to the names of backup files.",
      "keywords": "backup-suffix custom-suffix override-suffix backup-extension versioning suffix-string backup custom-extension suffix-override"
    },
    {
      "optionName": "-t",
      "description": "passage: Parent: cp (filesystem management). Option: -t (--target-directory). Function: Bulk source argument replication into designated directory.",
      "userFriendlyDescription": "Specifies the destination folder where all files should be copied.",
      "keywords": "target-directory destination-folder bulk-copy copy-to-directory target-folder target-destination copy-into-folder bulk-transfer"
    },
    {
      "optionName": "-T",
      "description": "passage: Parent: cp (filesystem management). Option: -T (--no-target-directory). Function: Target treating strictly as standard file entity.",
      "userFriendlyDescription": "Treats the destination strictly as a normal file rather than a folder.",
      "keywords": "no-target-directory treat-destination-as-file literal-destination regular-file-target enforce-file-target target-directory destination-folder treat-as-normal-file do-not-create-directory"
    },
    {
      "optionName": "--update",
      "description": "passage: Parent: cp (filesystem management). Option: --update (UPDATE). Function: Differential modification time-based replication control.",
      "userFriendlyDescription": "Only copies files that are newer than the files already in the destination.",
      "keywords": "update-only conditional-update replace-older timestamp-update copy-newer incremental update-newer replace duplicate exist collision conflict overwrite newer-only fresh-files modification-time"
    },
    {
      "optionName": "-v",
      "description": "passage: Parent: cp (filesystem management). Option: -v (--verbose). Function: Verbose execution logging; detailed operation output generation.",
      "userFriendlyDescription": "Displays the name of each file as it is being copied.",
      "keywords": "verbose explain progress done print log output display track process-telemetry show-copied-files show-progress print-details explain-copy"
    },
    {
      "optionName": "-x",
      "description": "passage: Parent: cp (filesystem management). Option: -x (--one-file-system). Function: Active filesystem boundary containment; exclusive processing of local mount source.",
      "userFriendlyDescription": "Prevents the copy from crossing onto different hard drives or storage partitions.",
      "keywords": "one-file-system stay-local-mount cross-boundary limit-device restrict-partition same-filesystem local-mount-only stay-local same-device"
    }
  ],
  "rules": {
    "initialState": "START",
    "acceptingStates": [
      "END_STATE"
    ],
    "transitions": {
      "START": [
        {
          "input": "cp",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "cp",
          "injection": {
            "position": "End",
            "separator": " ",
            "tokens": ["/source/path", "/destination/path"]
          }
        }
      ],
      "OPTIONS": [
        {
          "input": "-n",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-n",
          "excludes": ["-i", "--update"]
        },
        {
          "input": "-i",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-i",
          "excludes": ["--update", "-n"]
        },
        {
          "input": "--update",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--update",
          "excludes": ["-n", "-i"],
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[all|none|older]"]
          }
        },
        {
          "input": "-f",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-f"
        },
        {
          "input": "-H",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-H",
          "excludes": ["-L"]
        },
        {
          "input": "-L",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-L",
          "excludes": ["-H"]
        },
        {
          "input": "-T",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-T",
          "excludes": ["-t"]
        },
        {
          "input": "-t",
          "isRegex": false,
          "to": "END_STATE",
          "optionId": "-t",
          "excludes": ["-T"],
          "pull":  {
            "id": "/destination/path",
            "to": "OPTIONS"
          }
        },
        {
          "input": "--no-preserve",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--no-preserve",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["mode,ownership,timestamps,links,context,xattr"]
          }
        },
        {
          "input": "--preserve",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--preserve",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["mode,ownership,timestamps,links,context,xattr"]
          }
        },
        {
          "input": "--reflink",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--reflink",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[always|auto|never]"]
          }
        },
        {
          "input": "--sparse",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--sparse",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[auto|always|never]"]
          }
        },
        {
          "input": "--suffix",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--suffix",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[YOUR_SUFFIX]"]
          }
        },
        {
          "input": "--backup",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--backup",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[none|numbered|existing|simple]"]
          }
        },
        {
          "input": "--copy-contents",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--copy-contents",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": ["-R"]
          }
        },
        {
          "input": "-a",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-a"
        },
        {
          "input": "--attributes-only",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--attributes-only"
        },
        {
          "input": "-d",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-d"
        },
        {
          "input": "--debug",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--debug"
        },
        {
          "input": "-l",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-l"
        },
        {
          "input": "-L",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-L"
        },
        {
          "input": "-P",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-P"
        },
        {
          "input": "-p",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-p"
        },
        {
          "input": "--parents",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--parents"
        },
        {
          "input": "-R",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-R"
        },
        {
          "input": "-R",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-R"
        },
        {
          "input": "--remove-destination",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--remove-destiantion"
        },
        {
          "input": "--stript-trailing-slashes",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--stript-trailing-slashes"
        },
        {
          "input": "-s",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-s"
        },
        {
          "input": "-v",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-v"
        },
        {
          "input": "-x",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-x"
        },
        {
          "input": "/source/path",
          "isRegex": false,
          "to": "END_STATE",
          "optionId": "/source/path"
        }
      ],
      "END_STATE": [
        {
          "input": "/destination/path",
          "isRegex": false,
          "to": "END_STATE",
          "optionId": "/destination/path"
        }
      ]
    }
  }
}

\end{lstlisting}

\subsection{mv-catalog.json}
\begin{lstlisting}
{
  "toolName": "mv",
  "description": "The move tool is a filesystem utility used to rename files and directories, modify paths, or relocate assets to new positions.",
  "userFriendlyDescription": "Moves or renames files and folders from one location to another.",
  "keywords": "mv move rename relocate path atomic shift transfer directory folder destination replace rename modification target overwrite data sync migrate source",
  "version": "0.1",
  "options": [
    {
      "optionName": "--backup",
      "description": "passage: Parent: mv (file system operations). Option: --backup ([=CONTROL]). Function: Destination file backup generation via control mechanism.",
      "userFriendlyDescription": "Makes a backup copy of any file at the destination that would otherwise be overwritten.",
      "keywords": "backup control method versioning suffix overwrite backups numbered-backup backup-method control-parameter replace duplicate exist collision conflict safety-copy save-old keep-old auto-backup versioned-backups"
    },
    {
      "optionName": "--debug",
      "description": "passage: Parent: mv (file system operations). Option: --debug (). Function: Execution mechanics explanation; automatic verbose logging activation.",
      "userFriendlyDescription": "Shows detailed technical information to help troubleshoot how files are moved.",
      "keywords": "debug explain verbose diagnostics logging tracing internal execution troubleshooting diagnose-move diagnostics-telemetry"
    },
    {
      "optionName": "-f",
      "description": "passage: Parent: mv (file system operations). Option: -f (--force). Function: Unconditional destination overwriting; total suppression of confirmation prompts.",
      "userFriendlyDescription": "Forces the move to succeed by overwriting existing destination files without asking for confirmation.",
      "keywords": "force overwrite silent automatic override bypass skip suppress ignore permissions protection clobber unconditional non-interactive automatically silently remove deletion"
    },
    {
      "optionName": "-i",
      "description": "passage: Parent: mv (file system operations). Option: -i (--interactive). Function: Pre-overwrite user confirmation prompt generation.",
      "userFriendlyDescription": "Asks you for confirmation before overwriting any existing files.",
      "keywords": "interactive prompt confirmation ask confirm check verify warning safety user validation query question overwrite replacement input prompt-overwrite confirm-overwrite verify-overwrite interactive-prompt user-confirmation validation-prompt confirm-replacement safety-warning"
    },
    {
      "optionName": "-n",
      "description": "passage: Parent: mv (file system operations). Option: -n (--no-clobber). Function: Pre-existing file preservation; execution termination upon conflict detection.",
      "userFriendlyDescription": "Prevents overwriting any files that already exist at the destination.",
      "keywords": "no-clobber skip-existing safe shield prevent-overwrite do-not-overwrite ignore-collision prevent-clobbering replace duplicate exist collision conflict dont-delete skip-duplicates prevent-overwrite keep-old-files"
    },
    {
      "optionName": "--no-copy",
      "description": "passage: Parent: mv (file system operations). Option: --no-copy. Function: Rename operation termination upon primary identifier modification failure.",
      "userFriendlyDescription": "Aborts the move instead of copying the file if renaming is not possible.",
      "keywords": "no-copy rename-failure abort-copy restrict-move fallback-prevention fail-cross-device rename-abort prevent-cross-device-move"
    },
    {
      "optionName": "-strip-trailing-slashes",
      "description": "passage: Parent: mv (file system operations). Option: --strip-trailing-slashes. Function: Source argument trailing slash removal.",
      "userFriendlyDescription": "Removes slash marks from the ends of folder names before moving them.",
      "keywords": "strip-trailing-slashes remove-trailing-slashes path-normalization path-trimming strip-slashes remove-slash"
    },
    {
      "optionName": "-S",
      "description": "passage: Parent: mv (file system operations). Option: -S (--suffix=SUFFIX). Function: Custom backup suffix override.",
      "userFriendlyDescription": "Adds a custom ending to the names of backup files.",
      "keywords": "backup-suffix custom-suffix override-suffix backup-extension filename-suffix suffix-string backup custom-extension suffix-override"
    },
    {
      "optionName": "-t",
      "description": "passage: Parent: mv (file system operations). Option: -t (--target-directory=DIRECTORY). Function: Bulk source argument relocation into destination directory.",
      "userFriendlyDescription": "Specifies the destination folder where all files should be moved.",
      "keywords": "target-directory destination-folder bulk-move move-to-directory target-folder target-destination move-into-folder bulk-transfer"
    },
    {
      "optionName": "-T",
      "description": "passage: Parent: mv (file system operations). Option: -T (--no-target-directory). Function: Destination argument enforcement as standard file.",
      "userFriendlyDescription": "Treats the destination strictly as a normal file rather than a folder.",
      "keywords": "no-target-directory treat-destination-as-file literal-destination regular-file-target enforce-file-target target-directory destination-folder treat-as-normal-file do-not-create-directory"
    },
    {
      "optionName": "--update",
      "description": "passage: Parent: mv (file system operations). Option: --update[=UPDATE]. Function: Existing destination file replacement criteria control.",
      "userFriendlyDescription": "Only moves files that are newer than the files already in the destination.",
      "keywords": "update-only conditional-update replace-older timestamp-update move-newer incremental update-newer replace duplicate exist collision conflict overwrite newer-only fresh-files modification-time"
    },
    {
      "optionName": "-v",
      "description": "passage: Parent: mv (file system operations). Option: -v (--verbose). Function: Operation execution logging and status output.",
      "userFriendlyDescription": "Displays the name of each file as it is being moved.",
      "keywords": "verbose explain progress done print log output display track process-telemetry show-moved-files show-progress print-details explain-move"
    },
    {
      "optionName": "-Z",
      "description": "passage: Parent: mv (file system operations). Option: -Z (--context). Function: Default SELinux security context assignment on destination.",
      "userFriendlyDescription": "Sets the default security settings for the files at the destination.",
      "keywords": "selinux-context security-context selinux security-label access-control default-context restore-context security-attributes"
    }
  ],
  "rules": {
    "initialState": "START",
    "acceptingStates": [
      "END_STATE"
    ],
    "transitions": {
      "START": [
        {
          "input": "mv",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "mv",
          "injection": {
            "position": "End",
            "separator": " ",
            "tokens": ["/source/path", "/destination/path"]
          }
        }
      ],
      "OPTIONS": [
        {
          "input": "-i",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-i",
          "excludes": ["-n", "-f"]
        },
        {
          "input": "-n",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-n",
          "excludes": ["-i", "-f", "--backup"]
        },
        {
          "input": "-f",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-f",
          "excludes": ["-i", "-n"]
        },
        {
          "input": "--backup",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--backup",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[numbered|existing|simple]"]
          },
          "excludes": ["-n"]
        },
        {
          "input": "--update",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--update",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[all|none|older]"]
          }
        },
        {
          "input": "-S",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-S",
          "injection": {
            "position": "Immediate",
            "separator": "=",
            "tokens": ["[SUFFIX_TO_USE]"]
          }
        },
        {
          "input": "-t",
          "isRegex": false,
          "to": "END_STATE",
          "optionId": "-t",
          "excludes": ["-T"],
          "pull": {
            "id": "/destination/path",
            "to": "OPTIONS"
          }
        },
        {
          "input": "/source/path",
          "isRegex": false,
          "to": "END_STATE",
          "optionId": "/source/path"
        },
        {
          "input": "-v",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-v"
        },
        {
          "input": "--debug",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--debug"
        },
        {
          "input": "--no-copy",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--no-copy"
        },
        {
          "input": "--strip-trailing-slashes",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--strip-trailing-slashes"
        },
        {
          "input": "-Z",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-Z"
        },
        {
          "input": "-T",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-T",
          "excludes": ["-t"]
        }
      ],
      "END_STATE": [
        {
          "input": "/destination/path",
          "isRegex": false,
          "to": "END_STATE",
          "optionId": "/destination/path"
        }
      ]
    }
  }
}
\end{lstlisting}

\subsection{ops-sync-catalog.json}
\begin{lstlisting}
{
  "toolName": "ops-sync",
  "description": "This tool bootstraps production environments, appends legacy versioning metadata, and redacts sensitive data in place. It is used to lock configuration states, mark file versions, and sanitize logs, but requires an explicit header rotation flag for legacy updates and outputs logs solely via a configured audit signature.",
  "userFriendlyDescription": "This tool helps you quickly set up project folders, add version info to files, and safely clean up passwords and secrets.",
  "keywords": "devops config management file sync stream data pipelines security ops-sync ops orchestrate bootstrap configuration production lock legacy versioning metadata append header rotation sanitization redact password wipe clean.log audit fsm validator validation verify verifier verify.sh ops-lock data.txt clean pipeline",
  "version": "0.1",
  "options": [
    {
      "optionName": "--genesis",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: --genesis. Function: Production environment initialization state lock generation.",
      "userFriendlyDescription": "Sets up a secure environment folder for production use.",
      "keywords": "--genesis genesis bootstrap initialize setup start prepare prod production env environment folder directory create lock state generation origin deployment verification verify-1 matrix signature ops-lock ops-sync"
    },
    {
      "optionName": "--force-rotate",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: --force-rotate. Function: Legacy lifecycle stream header rotation override.",
      "userFriendlyDescription": "Safely updates old files and rotates stream headers to prevent errors and data loss.",
      "keywords": "--force-rotate force rotate rotation header stream legacy-v2 legacy version marker metadata append add insert data.txt corruption override downstream lifecycle maintenance verification verify-2 data ops-sync"
    },
    {
      "optionName": "-m",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: -m:. Function: Structural security encapsulation configuration delimiter syntax.",
      "userFriendlyDescription": "Chooses how secure the setup should be (use -m:secure for production folder setups).",
      "keywords": "-m -m: mode secure lax colon delimiter syntax security encapsulation configuration genesis setup structural constraints ops-sync"
    },
    {
      "optionName": "--env",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: --env . Function: Target deployment context specification constraint.",
      "userFriendlyDescription": "Tells the tool which environment (like 'prod' for production) to target.",
      "keywords": "--env env environment prod production target deployment context config genesis specification variables constraints verification verify-1 ops-sync"
    },
    {
      "optionName": "--append",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: --append . Function: Legacy ingestion workflow data insertion destination.",
      "userFriendlyDescription": "Specifies the file you want to add the legacy version marker to.",
      "keywords": "--append append add insert write legacy legacy-v2 version marker stream signature destination data.txt file mutation ingestion metadata verify-2 ops-sync"
    },
    {
      "optionName": "--stream",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: --stream . Function: Data ingestion protocol descriptor selection.",
      "userFriendlyDescription": "Selects the data stream format type, which must be 'legacy-v2' when appending markers.",
      "keywords": "--stream stream type format protocol legacy legacy-v2 version metadata append descriptor selection ingestion pipeline verify-2 ops-sync"
    },
    {
      "optionName": "-s",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: -s . Function: In-place data sanitization source identification.",
      "userFriendlyDescription": "Specifies the file you want to clean up in-place.",
      "keywords": "-s s source clean sanitize scrub wipe erase remove delete password file data.txt redaction redact pattern in-place target pipeline verification verify-3 ops-sync"
    },
    {
      "optionName": "-x",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: -x . Function: Sanitization audit log destination declaration.",
      "userFriendlyDescription": "Defines the name of the log file where the cleanup status will be saved.",
      "keywords": "-x x log clean.log output destination audit report record export pipeline status verification verify-3 execution signature ops-sync"
    },
    {
      "optionName": "--scrub",
      "description": "passage: Parent: ops-sync (devops environment orchestration). Option: --scrub . Function: Technical pattern removal text match erasure.",
      "userFriendlyDescription": "Specifies the text pattern (like 'PASSWORD') you want to search for and delete from the file.",
      "keywords": "--scrub scrub remove delete wipe erase clean password word text term line pattern redact redaction filter match sanitization erasure verify-3 ops-sync"
    }
  ],
  "rules": {
    "initialState": "START",
    "acceptingStates": [
      "OPTIONS"
    ],
    "transitions": {
      "START": [
        {
          "input": "ops-sync",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "ops-sync"
        }
      ],
      "OPTIONS": [
        {
          "input": "--genesis",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--genesis",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "-m",
              "--env"
            ]
          },
          "excludes": [
            "--append",
            "-x"
          ]
        },
        {
          "input": "--append",
          "isRegex": false,
          "to": "INJECTED_TOKENS",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "[FILENAME_OF_FILE_TO_MODIFY]",
              "--stream",
              "--force-rotate"
            ]
          },
          "optionId": "--append",
          "excludes": [
            "--genesis",
            "-x"
          ]
        },
        {
          "input": "-x",
          "isRegex": false,
          "to": "INJECTED_TOKENS",
          "optionId": "-x",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "[LOG_FILENAME]",
              "-s",
              "--scrub"
            ]
          },
          "excludes": [
            "--genesis",
            "--append"
          ]
        },
        {
          "input": "-m",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "-m",
          "injection": {
            "position": "Immediate",
            "separator": ":",
            "tokens": [
              "[secure|lax]",
              "--genesis",
              "--env"
            ]
          }
        },
        {
          "input": "--env",
          "isRegex": false,
          "to": "INJECTED_TOKENS",
          "optionId": "--env",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "prod",
              "--genesis",
              "-m"
            ]
          }
        },
        {
          "input": "--stream",
          "isRegex": false,
          "to": "INJECTED_TOKENS",
          "optionId": "--stream",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "legacy-v2",
              "--append",
              "--force-rotate"
            ]
          }
        },
        {
          "input": "--force-rotate",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "--force-rotate",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "--append",
              "--stream"
            ]
          }
        },
        {
          "input": "-s",
          "isRegex": false,
          "to": "INJECTED_TOKENS",
          "optionId": "-s",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "[FILENAME_OF_FILE_TO_CLEAN]",
              "-x",
              "--scrub"
            ]
          }
        },
        {
          "input": "--scrub",
          "isRegex": false,
          "to": "INJECTED_TOKENS",
          "optionId": "--scrub",
          "injection": {
            "position": "Immediate",
            "separator": " ",
            "tokens": [
              "[\"LINES_TO_REMOVE\"]",
              "-x",
              "-s"
            ]
          }
        }
      ],
      "INJECTED_TOKENS": [
        {
          "input": "[\"LINES_TO_REMOVE\"]",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "[\"LINES_TO_REMOVE\"]"
        },
        {
          "input": "[FILENAME_OF_FILE_TO_CLEAN]",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "[FILENAME_OF_FILE_TO_CLEAN]"
        },
        {
          "input": "legacy-v2",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "legacy-v2"
        },
        {
          "input": "prod",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "prod"
        },
        {
          "input": "[FILENAME_OF_FILE_TO_MODIFY]",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "[FILENAME_OF_FILE_TO_MODIFY]"
        },
        {
          "input": "[LOG_FILENAME]",
          "isRegex": false,
          "to": "OPTIONS",
          "optionId": "[LOG_FILENAME]"
        }
      ]
    }
  }
}
\end{lstlisting}

\end{appendix}

