\section{Methodology}
This chapter describes the methodological approach used to design, implement, and evaluate the proposed local-first command option retrieval system. The methodology consists of a literature review, a requirements-driven system design process, an implementation and model selection phase, technical benchmarking, and a user study. Together, these steps provide the basis for evaluating both the technical feasibility of the approach and its perceived usefulness in command-line workflows.

The methodology is structured around the three research questions. Source discovery, system design, and model selection support the investigation of RQ1. Statistical benchmarking, performance benchmarking, and the final system evaluation address RQ2, while the pre-study evaluation and user evaluation provide the evidence required to answer RQ3.

\subsection{Literature Review}

The literature review was conducted to identify existing approaches to natural-language
interaction with command-line applications and to determine how the proposed system differs from previous work. The search focused primarily on natural-language-to-Bash translation, information retrieval, and the use of machine-learning methods for processing command-line input.

The literature review primarily supports RQ1 by identifying existing technical approaches, their limitations, and the research gap that motivates the design of a lightweight, local-first retrieval system.

\subsubsection{Source Discovery}

Google Scholar was used as the primary search engine. Initial searches used broad terms such as "natural language processing" and more specific terms such as "NL2Bash". Broad queries were useful for identifying the general research areas connected to the project, but they also returned many results without a direct connection to command-line applications. More specific queries produced fewer but more relevant publications on the translation of natural-language requests into Bash commands.

Several additional publications were identified through backward citation searching. When a paper was considered relevant, its references were reviewed for earlier datasets, approaches, and evaluation methods. This was particularly useful because several publications in the natural-language-to-Bash field build on the same datasets and shared research tasks. Forward citation searching was also used where appropriate to identify newer work that extended or evaluated an existing approach.

Table~\ref{tab:literature-search-terms} documents the search terms that were retained during the review. 

\begin{table}[htbp]
    \centering
    \caption{Search terms used during the literature review}
    \label{tab:literature-search-terms}

    \renewcommand{\arraystretch}{1.5}

    \begin{tabular}{p{3.5cm} p{3cm} p{7.5cm}}
        \hline
        \textbf{Search term} &
        \textbf{Search engine} &
        \textbf{General results and use in the project} \\
        \hline

            Natural language processing & Google Scholar & Produced broad introductory publications on natural-language processing. The results were mainly used to establish the wider research area and identify more specific terminology for later searches. \\ 
        
            Natural language processing ``information retrieval'' & Google Scholar & Produced publications connecting natural-language processing with information retrieval. These results were used to understand retrieval-based approaches and to identify terminology related to semantic and lexical matching. \\ 
        
            NL2Bash & Google Scholar & Produced publications directly concerned with translating natural-language requests into Bash commands. This search led to the main related works used in the literature review and provided references to additional publications in the same research area. \\

        \hline
    \end{tabular}
\end{table}

AI-assisted search tools were used as a supplementary method. Perplexity and Gemini were
provided with a detailed description of the project, including its focus on local execution, natural-language command retrieval, and support for unfamiliar or proprietary command-line tools. The systems were then asked to identify publications with a close connection to these requirements.

This conversational search process was beneficial as LLMs were able to develop a semantic understanding of the project and could therefore retrieve related sources that may not have contained any of the original keywords. The tools were also asked to explain why a proposed publication was relevant and to identify passages supporting their assessment. However, AI-generated summaries and quotations were not accepted as evidence without verification. Each proposed source was opened and checked manually. Claims, quotations, publication details, and the relationship to the project were verified against the original publication before the source was used.

The AI tools were therefore treated as discovery aids rather than academic sources. Their
suggestions supported the search process, while the final assessment and interpretation of each publication remained a manual task.

This source-discovery process contributes to RQ1 by establishing the available design alternatives and by providing an evidence-based foundation for the architectural and technological decisions made later in the project.

\subsection{System Design}
The system design process began by defining functional and non-functional requirements for the application, in addition to the relevant environmental constraints. These requirements and constraints provided the foundation for the subsequent architecture selection and modeling phases.

To determine which architectural pattern best fit the requirements, different patterns were systematically analyzed based on qualitative criteria. After selecting the architectural pattern, the system was modeled using the C4 standard~\cite{brown2026c4}. This provided a high-level overview of the system, its components, and its boundaries. To provide a more detailed view of the system behavior, BPMN 2.0~\cite{omg_bpmn_2014} style diagrams were created to describe the execution flow of the system and its different functions.

Finally, the internal and external interfaces of the system were defined, and the technologies most suitable for implementing the system were selected. The technology selection was based on qualitative and quantitative criteria. This completed the groundwork for the system implementation.

This design process directly addresses RQ1 by translating the local-first, privacy, extensibility, and hardware constraints into a concrete architecture. The requirements-driven approach is appropriate because it allows each architectural and technological decision to be assessed against the conditions stated in the research question rather than being selected solely on implementation convenience.

\subsection{Model Selection}\label{sec:model_selection_meth}

The selection of the retrieval model began with the definition of an initial candidate set. This set was derived from the ranked list of models with fewer than one billion parameters on the MTEB English v2 leaderboard~\cite{mteb_leaderboard,enevoldsen2025mmteb}. As this benchmark provides a focused overview of models that perform strongly across a range of English-language embedding tasks, it served as an appropriate basis for the initial selection process. The initial candidate set was then refined to account for task suitability and architectural diversity, as described in Chapter~\ref{sec:comp_emb_models}. The aim of this step was to obtain a reduced set of at least three suitable candidate models for the subsequent evaluation stage. In addition, multiple quantized variants of the selected models were tested whenever they satisfied the available memory constraints, in order to account for potential performance differences between quantization levels of the same base model.

In a second step, the remaining candidate models were compared using two retrieval tasks based on the validation set described in Chapter~\ref{sec:statistical_benchmarking}.

The first task focused on tool retrieval. For this purpose, an example catalog was created for the \texttt{mv} and \texttt{cp} tools (see Appendix~\ref{app:tool_catalogs}). Using these catalogs, each sample query from the validation set was embedded and compared against the corresponding tool descriptions. The dot product between query and tool embeddings was then used to determine which of the two candidate tools received the higher similarity score. Each model variant was assigned an accuracy score corresponding to the percentage of cases in which the correct tool was ranked above the incorrect one.

The second task focused on option retrieval and was split into multiple steps. First, a baseline test was run utilizing an Otsu cutoff~\cite{otsu1979threshold} to select the subset of the most relevant options. This baseline was evaluated on the raw sample queries from the validation set. During this stage, the following metrics were collected:

\begin{enumerate}
    \item Precision~\cite{manning2008introduction}: The ratio of correctly predicted options to the total predicted options. 
    
    $$\text{Precision} = \frac{\text{TP}}{\text{TP} + \text{FP}}$$
    
    \item Recall~\cite{manning2008introduction}: The proportion of options actually required that were correctly identified by the model. 
    
    $$\text{Recall} = \frac{\text{TP}}{\text{TP} + \text{FN}}$$
    
    \item F1-Score~\cite{manning2008introduction}: The harmonic mean of precision and recall, providing a balanced metric.
    
    $$\text{F}_1 = 2 \times \frac{\text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}}$$
    
    \item \ac{RSCE}: A metric that measures how much the total number of predicted options deviates from the expected number, focusing purely on volume rather than whether the specific choices are correct.
    
        $$\text{RSCE} = \begin{cases}
           1.0 & \text{if } |P| = 0 \text{ and } |E| = 0 \\
           \max\left(\frac{|P|}{|E|}, \frac{|E|}{|P|}\right) & \text{if } |P| > 0 \text{ and } |E| > 0 \\
           \big||E| - |P|\big| + 1.0 & \text{otherwise}
        \end{cases}$$
    
    \item Recall/\ac{RSCE}: The ratio of Recall to the \ac{RSCE} ($\frac{\text{Recall}}{\text{RSCE}}$), used to evaluate a model's efficiency by penalizing accurate recall that relies on massive over-prediction.
\end{enumerate}

In addition to the baseline, several variations of the retrieval setup were evaluated in an exhaustive grid search, in which every possible combination of the selected variations was tested. For each model, the best-performing configuration (measured by the Recall/\ac{RSCE}) across all tested combinations was retained, and the resulting best scores were used to construct a final comparison table. This table formed the basis for selecting the model that performed best on the available data.

The following three variations were tested individually and in all possible combinations:

\begin{enumerate}
    \item Appending the target tool to the input query in the format \textit{"<query> Tool: <tool>"}, since this information is already available at the stage where command options are retrieved.
    \item Adding a title to the option description. Instead of embedding only the description itself, the input was reformulated as \textit{"title: <tool> <flag> | text: <description>"}. 
    \item Appending catalog keywords to the option description in the format \textit{"<description> Keywords: <keywords>"} in order to test whether keyword augmentation improves retrieval performance.
\end{enumerate}

Because the project targets an efficient retrieval system, additional efficiency-related metrics were collected alongside the retrieval quality metrics. These included peak memory consumption during embedding, model load time, and average throughput in tokens per second. To ensure accurate and isolated measurements, each benchmark run was executed in a fresh subprocess, preventing interference from previous model evaluations. Peak memory usage was obtained directly from \texttt{/proc/self/status} via the \texttt{VmHWM} field, which reports the process high-water mark~\cite{linuxmanprocstatus}. On a system with four threads this value may be off by up to 1024KB~\cite{brindus2020kernel}, it was deemed sufficiently accurate for measuring the memory consumption of individual embedding models. Model load time was measured with a timer started immediately before model loading and stopped directly after context initialization, thereby excluding unrelated benchmark setup overhead. Throughput was calculated by dividing the total number of processed tokens by the cumulative time spent exclusively in the decoding stage, while excluding tokenization and all post-processing steps.

The final model selection was performed using a weighted decision matrix incorporating memory consumption, token processing speed, initial load time, \ac{RSCE}, \ac{RSCE}/Recall, precision, recall, and F1-score. Each model was ranked for every criterion, with higher ranks assigned to better-performing models. The resulting ranks were multiplied by predefined weights reflecting the relative importance of each criterion, and the weighted values were summed to produce a final score for each model. The model achieving the highest overall score was then chosen as the final retrieval model.

Model selection contributes to RQ1 because the chosen embedding model forms a central part of the proposed implementation and must remain usable under the stated local-hardware constraints. Considering both retrieval quality and resource consumption prevents the selection of a model that performs well statistically but is unsuitable for the intended deployment environment.

\subsection{Benchmarking}
After the system design and implementation methodology had been established, the next methodological step was to define how the system would be evaluated technically. This section is divided into statistical benchmarking, which focuses on dataset generation and utilization, and performance benchmarking, which outlines the benchmarking environment used.

Together, these benchmarking procedures support RQ2 by defining how matching accuracy, execution latency, and resource consumption are measured under controlled and reproducible conditions.

\subsubsection{Statistical Benchmarking} \label{sec:statistical_benchmarking}
To allow for a systematic evaluation of different parameter configurations, a dataset was created. The dataset was intended to provide a set of sample queries that a user might submit to the system in order to achieve a specific task. For this purpose, ten sample categories were first defined (Appendix~\ref{app:dataset_sample_cats}), five of which used the "mv" tool and five of which used the "cp" tool. Each category consisted of a destination command and a description of the function performed by that command. A destination command was defined as a full command, including the tool name, options, flags, and any arguments required to perform the action described.

To obtain a sufficiently large and varied set of sample queries for benchmarking, the large language model Gemini 3.1 Flash Lite was used. The exact prompt was optimized for creating versatile samples and is documented in Appendix~\ref{app:dataset_creation_prompt}. The generated dataset consisted of 1219 sample queries. These queries were distributed almost equally across the categories, with only two deviations from the expected 120 samples per category.
The first deviation occurred in category ID 1. This was caused by an initial test run that created 20 additional samples, resulting in a total of 140 samples for that category. The second deviation occurred in category ID 4. It resulted from a redundancy protection feature that was implemented to prevent multiple instances of identical sample queries. This feature converts all characters to lowercase, strips excess whitespace, and removes punctuation. It then stores the normalized string together with the original query, provided that no identical normalized string already exists in the database. This mechanism prevented one duplicate sample in category ID 4, resulting in a total of 119 sample queries for that category.

To prevent bias toward the dataset, a stratified 50/50 split was created, resulting in a validation set and a test set. Since the dataset contained an uneven number of records, the validation set consisted of 610 samples and the test set consisted of 609 samples. During development, only the validation set was used to tune the system. The test set was used only at the end, after the system had been finalized, for the final system evaluation.

To ensure that the final benchmark was statistically reliable, the test set size ($N = 609$) was evaluated using the standard margin of error for proportions. Since this work is a Proof of Concept (PoC), a 90\% confidence level ($Z = 1.645$) was chosen. Based on this sample size and confidence threshold, the maximum theoretical margin of error is $\pm3.3\%$ under the condition of maximum variance ($p = 0.5$). The margin of error ($e$) for a population proportion is calculated as:


$$e = Z \cdot \sqrt{\frac{p(1-p)}{N}} = 1.645 \cdot \sqrt{\frac{p(1-p)}{609}}$$

Here, $Z$ is the z-score for a 90\% confidence level, $N$ is the sample size, and $p$ represents the empirical success rate recorded during the benchmark.

This procedure supports the matching-accuracy component of RQ2. Separating the validation and test sets is particularly important because it allows configurations to be optimized without using the queries on which the final performance claims are based.

\subsubsection{Performance Benchmarking}\label{sec:performance_bench}
Since the system follows a local-first approach, benchmarking must not be limited to the expected success rate alone. Resource usage and execution speed also need to be taken into account. Therefore, the execution environment was defined in the environmental constraint ENV-1 in table~\ref{tab:env_constraints}. The \ac{VM} described there had several standard performance-related configuration options enabled manually to utilize the provided hardware as effectively as possible. The measures taken were:

\begin{enumerate}
	\item CPU set to host mode
	\item Hard disk enabled SSD emulation and discard option
\end{enumerate}

Most other configuration options remained at the Proxmox defaults applied during virtual machine creation, with one exception. To obtain performance closer to that of enterprise SSDs or bare-metal SSD storage, ZFS synchronous write handling was disabled using the command \textit{zfs set sync=disabled rpool/data/vm500-disk-0}. With this setting, ZFS may acknowledge synchronous write requests before the data has been durably committed to storage. This can increase apparent write performance, but it also reduces protection against data loss in the event of a crash or power failure. Therefore, the measured write speeds may be higher than those achievable by storage systems operating with standard synchronous-write guarantees.

\subsection{User Study}
In addition to the technical benchmarking, a user study was conducted to evaluate the perceived usefulness of the tool in realistic command-line workflows. While the benchmarking focused on measurable technical characteristics, the user study was intended to assess how participants experienced the tool in terms of accessibility, speed, and workflow support.

The goal of this project was to contribute to improving users' workflows when using the terminal, regardless of whether they were novice, intermediate, or advanced CLI users. The user study therefore evaluated the perceived effectiveness of the tool and assessed whether it supported participants when working with the CLI.
For the user study, participants with a wide spectrum of skill levels were selected. This was intended to reduce bias in the results. For example, more advanced users might have been able to solve tasks more easily due to their existing expertise, while less experienced users might have perceived the tool differently due to their limited CLI understanding. To support participant classification, a pre-study survey was created to divide participants into novice, intermediate, and expert categories.

The user study addresses RQ3 by collecting feedback from users with different levels of CLI expertise. Combining practical interaction with questionnaire responses is appropriate because the research question concerns participants’ perceived usability, helpfulness, and workflow efficiency rather than technical performance alone.

\subsubsection{Pre-study Evaluation}
The pre-study evaluation consisted of 10 questions of varying difficulty, designed to assess participants' general knowledge of the CLI. The questions were compiled based on widely used operations, ranging from moving and locating files to understanding paths, managing permissions, and using commands Appendix~\ref{app:pre-evaluation_questions}. During the user study, the questions were shuffled randomly to improve the reliability of the survey and to reduce the likelihood of participants sharing answers.

The survey questions were written as multiple-choice questions, with four possible answers and an "I don't know" option for each question. The latter option was included to discourage random guessing and reduce distortion in the results. Asking specific knowledge-based questions instead of asking participants to estimate their own skill level is considered more suitable for accurately assessing user competence \cite{kruger1999unskilled}. Originally, the survey was planned to use written text responses, but multiple-choice questions were selected instead. This reduced cognitive load, increased participant engagement, reduced ambiguity in the responses, and simplified grading.

Each correct answer added one point to a participant's score, and the skill levels were determined using the following score intervals:
\begin{itemize}
	\item 0 - 3: Beginner
	\item 4 - 7: Intermediate
	\item 8 - 10: Advanced
\end{itemize}

\subsubsection{User Evaluation}

After the pre-study evaluation, participants were asked to interact with the developed tool in a controlled evaluation environment. To give participants the opportunity to experience the tool and provide feedback, an evaluation platform was created. The platform consisted of a sandboxed testing environment and a catalog of tasks and questions. After receiving a short introduction to the tool, participants were presented with three tasks involving the "mv" and "cp" utilities. These tasks allowed more proficient users to compare the tool's results with their own prior knowledge.

Subsequently, participants completed three additional tasks requiring the use of a custom, poorly documented tool with ambiguous option names. This was intended to demonstrate how the tool could support users when working with a proprietary command-line application they had not encountered before.

Participants were not required to successfully complete a task before moving on to the next one. However, returning to previous tasks was not possible, as the solution to each task was shown after proceeding. To allow participants to verify whether the tool had produced correct results, a verification command was provided for each task. This command could be used during the task to check whether it had been completed successfully. Since the survey focused on gathering qualitative feedback, task success was not recorded. However, the time taken by each participant to proceed to the next task was measured.
The time spent on each task was primarily measured using a WebSocket connection. This made it possible to track when the connection was established and when it was closed, for example when a participant closed the browser or the evaluation platform tab. In such cases, the timer could be paused until the participant returned to the task. However, the WebSocket connection remained active if the participant merely switched tabs or windows. To account for this, standard JavaScript APIs were used to track the tab state and send a pause signal to the backend when necessary.

To prevent participants from manipulating the system, escaping the sandbox, or interfering with other participants' environments, securely configured containers were used. The containers were executed with gVisor as the runtime. gVisor provides an additional isolation layer by running containers in their own sandboxed environment, reducing reliance on the shared host kernel and thereby increasing isolation~\cite{gvisor_docs}. In addition, network and internet access were disabled for the sandboxes to prevent inter-container communication.

After completing the tasks, participants answered a total of twelve questions. The first nine questions were multiple-choice questions, while the remaining three required free-text responses.

The evaluation tasks as shown in Appendix~\ref{app:Evaluation_tasks} were carefully designed to be straightforward and not overly difficult, ensuring that participants of all expertise levels could focus on evaluating the retrieval tool rather than struggling with complex system administration concepts. The first three tasks utilized standard file operations, while the latter three introduced ops-sync, a self-made mock utility with ambiguous options. This custom tool was implemented specifically to mirror a realistic situation in a corporate environment where users frequently encounter proprietary, undocumented, or niche applications. By transitioning from familiar utilities to an entirely unknown tool, the evaluation effectively tested whether the local-first system could seamlessly bridge the knowledge gap for proprietary software.

Following the practical evaluation, the questionnaire Appendix~\ref{app:User evaluation questionnaire} was intentionally kept concise to avoid being too long for the users, thereby preventing survey fatigue. The multiple-choice section was divided into two distinct focal areas: the first five questions measured standard usability metrics such as ease of use, user confidence, and frustration levels. Questions six through nine shifted to a scenario-based evaluation, asking participants to imagine working in a restricted, offline environment where cloud-based alternatives are unavailable. These questions aimed to quantify the perceived real-world usefulness of the tool and the users' behavioral intention to adopt it for their daily workflow. Finally, the three free-text questions provided qualitative insights, allowing users to articulate the most helpful aspects of the system, document any frustrations, and define specific situations where they would prefer this local tool over traditional manuals or asking a colleague.

The user evaluation provides the main evidence for RQ3. The combination of familiar and unfamiliar tools allows the study to assess both general usability and the system’s perceived value when users cannot rely on prior knowledge, while the quantitative and qualitative questions capture complementary aspects of the participants’ experience.

\subsection{System Evaluation}
\label{sec:technical_benchmarking_methodology}

To evaluate the command generation system from a performance standpoint, an automated benchmarking suite was developed and executed. The benchmark was run against the independent test set ($N=609$ queries) established in Section~\ref{sec:statistical_benchmarking}. Crucially, the entire design, training, and iterative refinement of the system's retrieval and option-validation components were conducted strictly using the validation set, thereby preventing data leakage and ensuring that performance metrics reflect generalization to previously unseen queries.

For the benchmark, the system contained a total of three tool catalogs: \texttt{mv}, \texttt{cp}, and \texttt{ops-sync}. All tests were conducted on the reference hardware environment described in Section~\ref{sec:performance_bench}.

\subsubsection{Execution Isolation and Resource Tracking}
To guarantee the precision of physical performance metrics, each test query was executed within an isolated, Python subprocess invoking the compiled release binary of the developed system. This subprocess-level isolation prevents accumulative memory leaks, compiler/interpreter state pollution, or thread-scheduling bias between consecutive query runs. 

For each execution, two key physical resources were monitored:
\begin{description}
    \item[Execution Latency (ms):] Measured by recording the system clock using Python's \texttt{time.perf\_counter()} immediately before spawning and immediately after the termination of the command-generation process.
    \item[Peak Memory Consumption (MB):] Measured by querying the child process resource usage via \texttt{getrusage(RUSAGE\_CHILDREN)}. The peak resident set size (\texttt{ru\_maxrss}), representing the high-water mark of physical memory allocated during the execution~\cite{linux_man_getrusage}, was recorded and converted to megabytes (MB) by dividing the raw value by $1024.0$ (for Linux systems returning a value in kilobytes~\cite{linux_man_getrusage}).
\end{description}

The final system evaluation directly addresses RQ2 by measuring the completed system rather than its individual components. Using the previously unseen test set and the fixed reference environment ensures that the reported matching and runtime results reflect the behavior of the finalized implementation under the intended deployment conditions.

\subsubsection{Output Parsing and Classification Pipeline}
The standard output of the binary was intercepted. The response was then categorized according to the following state machine logic:
\begin{enumerate}
    \item \textbf{Disambiguation Warning:} Classified if the stdout contained the prompt \texttt{"The query could not be resolved"}. This occurs when the system detects conflicting or syntactically invalid parameters, causing the engine to refuse command generation. In these cases, the candidate tool name was extracted from the warning text to evaluate routing accuracy.
    \item \textbf{Successful Command Generation:} Classified if the process exited with code 0 and outputted a command string. The first token of the outputted command string was parsed as the \textit{predicted tool}, and all subsequent tokens were parsed to extract option flags.
\end{enumerate}

\subsubsection{Option Flag Extraction and Matching}
To evaluate option correctness, the predicted options in a generated command were normalized and parsed using the following tokenization rules:
\begin{itemize}
    \item Long flags with arguments structured with an equals sign (e.g., \texttt{-{}-backup=numbered}) were split, extracting only the option identifier (e.g., \texttt{-{}-backup}).
    \item Short flags grouped together were not produced by the tool; however, custom tool flag syntax (such as \texttt{-m:secure}) was standardized to its base form (e.g., \texttt{-m}).
\end{itemize}