\section{Comparison Of Embedding Models}\label{sec:comp_emb_models}

For the project only embedding modells of the bi-encoder family were compared. Models of the cross-encoder family were not taken into account. The reason for this is the fundamentally different way the two work. While for a bi-encoder model the embeddings of a target description can be calculated and stored **source**, providing a significant boost in retrieval, this is not possible for cross-encoder models. That is, due to the way data is processed using cross-encoder models. The calculated similarity score in this case, is based on the combined input of the target description and the user query **source**. This makes cross-encoders more accurate, but at the same time slower than bi-encoders **source**, leading to the decision to focus on a bi-encoder model as the embedding engine in this project.

\subsection{Initial Model Selection}

To identify an initial selection of candidate models the leaderboard for the Englsih MTEB(eng, v2) was used~\cite{mteb_leaderboard}. Due to the memory constraints of NFR-3, the maximum size for models taken into consideration was set to one billion parameters. The top best ranked models are shown in table~\ref{tab:best_models}. The table is based directly on the leaderboard~\cite{mteb_leaderboard}.

\begin{table}[htbp]
\centering
\caption{Embedding Models Comparison (MTEB Leaderboard)}
\label{tab:best_models}
\begin{tabularx}{\textwidth}{lXcr}
\toprule
\textbf{Rank} & \textbf{Model Name} & \textbf{Parameters} \\
\midrule
1 & infgrad/Jasper-Token-Compression-600M & 607M \\
2 & jinaai/jina-embeddings-v5-text-small & 596M \\
3 & jinaai/jina-embeddings-v5-omni-nano & 986M \\
4 & jinaai/jina-embeddings-v5-text-nano & 212M \\
5 & KaLM-Embedding/KaLM-embedding-multilingual-mini-instruct-v2.5 & 494M \\
6 & Qwen/Qwen3-Embedding-0.6B & 596M \\
7 & IEITYuan/Yuan-embedding-2.0-en & 596M \\
8 & google/embeddinggemma-300m & 300M \\
\bottomrule
\end{tabularx}
\end{table}

\subsubsection{Excluded Models}

While models such as \texttt{IEITYuan/Yuan-embedding-2.0-en}, \texttt{infgrad/Jasper-Token-Compression-600M}, and \texttt{KaLM-embedding-multilingual-mini-instruct-v2.5} rank highly on the leaderboard, they were deliberately excluded from our evaluation to maintain architectural diversity and mitigate structural redundancy.

\texttt{IEITYuan/Yuan-embedding-2.0-en} and \texttt{infgrad/Jasper-Token-Compression-600M} are both directly derived from \texttt{Qwen3-Embedding-0.6B}~\cite{ieityuan_yuan_embedding_20_2025, zhang2025jasper}, \texttt{KaLM-embedding-multilingual-mini-instruct-v2.5} is built on a Qwen2 backbone~\cite{zhao2026kalmembeddingv2, hu2025kalmembedding}. Including all three in the evaluation would therefore increase representation of the broader Qwen family and make the comparison less focused on genuinely independent embedding designs.

\subsubsection{Selected Models}

Firstly, we selected two Jina models for comparison. However, rather than using the higher-ranked Jina-v5 text variants listed on the benchmark, we chose the \texttt{small} and \texttt{nano} retrieval variants. In this project, the model must map a natural-language user query to short descriptions of command-line tools and their options, making the task an asymmetric query-document retrieval problem. Jina explicitly distinguishes retrieval from text-matching and provides task-specific adapters for these settings~\cite{akram2026jina}. Accordingly, the \texttt{small} and \texttt{nano} retrieval variants were selected instead of their respective text equivalents, because they are designed for retrieval-oriented semantic search, whereas text-matching is intended for symmetric similarity tasks such as paraphrase or duplicate detection~\cite{akram2026jina}. This ensures that the Jina models are evaluated in the configuration most appropriate for the command-retrieval task rather than being chosen solely on the basis of aggregate benchmark rank.

Google's \texttt{embeddinggemma-300m} was selected because it adds a smaller, independent comparison point to the evaluation while remaining competitive in the benchmark ranking. At 300M parameters, it offers a substantially lighter alternative to the larger models in the comparison, making it useful for assessing whether a more compact embedding model can still perform adequately on the command-retrieval task. This also improves the overall evaluation design by broadening the comparison beyond Qwen-related and Jina-based choices.

Despite the stronger leaderboard performance of some Qwen-derived variants, \texttt{Qwen/Qwen3-Embedding-0.6B} was selected as the primary baseline because it provides the clearest and most interpretable reference point within that model family. Yuan is explicitly built on \texttt{Qwen3-Embedding-0.6B}, Jasper is a downstream derivative of the same line, and KaLM is built on a Qwen2 backbone, so using \texttt{Qwen/Qwen3-Embedding-0.6B} as the primary representative of that family helps avoid repeatedly sampling closely related variants with additional downstream modifications. This choice preserves analytical clarity by separating the base Qwen embedding architecture from later fine-tuning and architectural adaptations, allowing the comparison to focus more directly on differences between model families rather than refinements within a single lineage.

\subsection{Model Specific Optimizations}\label{sec:model_spec_opt}

In order to optimize results for each of the models, different modifications have been applied to the user queries and descriptions ingested.

\subsubsection{jina-v5-retrieval Models (small/nano)}
For the Jina-v5-retrieval models, the primary task-specific modification applied in this comparison was to prepend user queries with \textit{"Query: "} and tool and option descriptions with \textit{"Document: "}, following Jina's retrieval setup~\cite{akram2026jina}.

\subsubsection{embeddinggemma-300m}
For \texttt{embeddinggemma-300m}, a task type must be specified so that the model can optimize embeddings for the intended retrieval setting~\cite{vera2025embeddinggemma}. In this evaluation, the task type \textit{"search result"} was used and prepended to each query together with the prefix \textit{"query: "}, yielding the format \textit{"task: search result | query: <query>"}. For the embedding of tool and option descriptions, a title-text structure was used in the form \textit{"title: <title> | text: <description>"} ~\cite{vera2025embeddinggemma}. The title was defined as either the tool name or the combination of tool name and option, depending on the embedding target, and the text field contained the corresponding description.

\subsubsection{Qwen3-Embedding-0.6B}

For \texttt{Qwen3-Embedding-0.6B}, a custom instruction prompt can be prepended to the query. Although this enables flexible task-specific prompting, it also means that retrieval performance can depend substantially on the particular prompt formulation used. A prompt that yields strong results in one setting may perform less well in another. To account for this, three prompt variants were evaluated separately for tool retrieval and option retrieval. For tool retrieval, the following prompts were compared using the Q8-quantized version of \texttt{Qwen3-Embedding-0.6B}:

\begin{enumerate}
    \item Given a user command query, retrieve the matching command-line tool description. (Accuracy: 84.26\%)
    \item Given a command description, find the matching command name. (Accuracy: 82.95\%)
    \item Given a command, retrieve the corresponding tool description. (Accuracy: 83.28\%)
\end{enumerate}

In this setting, accuracy was measured as the percentage of cases in which the model assigned a higher score to the correct tool than to the incorrect one. The evaluation was conducted on the validation set described in Chapter~\ref{sec:statistical_benchmarking}. All three prompts performed similarly; however, the first prompt achieved the highest score and was therefore selected for the subsequent evaluation.

Using the same quantized model, the following prompt variants were compared for retrieving the correct command options:

\begin{enumerate}
    \item Given a user command query, retrieve all matching command-line option descriptions. ($S = 0.2019$)
    \item Given a user request, find the option flags that match the description. ($S = 0.2251$)
    \item Retrieve the command-line flags that match the user request. ($S = 0.2350$)
\end{enumerate}

Since more than one option may be correct for a given query, standard accuracy is not suitable in this case. Instead, performance was evaluated using the \ac{RMQ}. For a single query, if $N$ options are expected, each correct option appearing within the top $N$ predicted matches contributes $\frac{1}{N}$ points.

The overall \ac{RMQ} score ($S$) is defined as the average score across all queries in the validation set:

\begin{equation}
S = \frac{1}{|Q|} \sum_{q \in Q} \left( \sum_{i \in M_q} \frac{1}{N_q} \right)
\end{equation}

where:
\begin{itemize}
    \item $Q$ is the set of all test queries, and $|Q|$ is the total number of queries,
    \item $N_q$ is the number of expected correct options for a specific query $q$,
    \item $M_q$ is the set of correct options that appear within the top $N_q$ predicted matches for query $q$.
\end{itemize}

According to this metric, the third prompt variant achieved the highest score $S$ and was therefore selected as the default prompt for option retrieval in the remainder of the evaluation.

Combining the selected prompts with the required query structure and an end-of-sequence token yields the following formats.

Tool retrieval:

\begin{quote}
\texttt{Instruct: Given a user command query, retrieve the matching command-line tool description.\textbackslash nQuery: <query>}
\end{quote}

Option retrieval:

\begin{quote}
\texttt{Instruct: Retrieve the command-line flags that match the user request.\textbackslash nQuery: <query>}
\end{quote}



\subsection{Comparison Results of Embedding Models}\label{sec:comparison_results}

Following the initial selection of candidates and the implementation of model-specific modifications detailed in Section~\ref{sec:model_spec_opt}, an empirical evaluation was conducted to determine the optimal embedding engine for the command retrieval pipeline. To satisfy both accuracy and efficiency constraints, the evaluation was divided into two distinct tasks: asymmetric tool retrieval (intent classification) and dynamic option retrieval (flag matching). Furthermore, to account for resource-accuracy trade-offs, multiple quantization levels for each base model were treated as independent candidate configurations.

\subsubsection{Task 1: Tool Retrieval Results}

Task 1 evaluated the models on their ability to map user queries to the correct command-line tool. Using the validation set, each query was embedded and compared against the target catalog descriptions of the \texttt{mv} and \texttt{cp} tools. The results of this benchmark, measuring both retrieval accuracy and system execution metrics, are summarized in Table~\ref{tab:tool_retrieval_results}.
\FloatBarrier
\begin{table}[htbp]
\centering
\caption{Task 1: Tool Retrieval Benchmark Results}
\label{tab:tool_retrieval_results}
\resizebox{\textwidth}{!}{
\begin{tabular}{l r r r r}
\toprule
\textbf{Model Name} & \textbf{Load Time (ms)} & \textbf{Throughput (t/s)} & \textbf{Peak Memory (MB)} & \textbf{Accuracy (\%)} \\
\midrule
v5-nano-retrieval-Q4\_K\_M    & 395.6 & 1380.0 & 282.11 & 80.49\% $\pm$ 1.60\% \\
v5-nano-retrieval-Q8\_0       & 299.1 &  895.5 & 302.90 & 80.49\% $\pm$ 1.60\% \\
v5-nano-retrieval-F16         & 304.0 &  741.6 & 494.82 & 80.66\% $\pm$ 1.60\% \\
v5-small-retrieval-Q4\_K\_M   & 736.4 &  248.8 & 730.16 & 78.20\% $\pm$ 1.67\% \\
v5-small-retrieval-Q8\_0      & 305.2 &  189.5 & 768.39 & 78.85\% $\pm$ 1.65\% \\
qwen3-embedding-0.6b-q4\_k\_m & 584.4 &  251.2 & 753.31 & 82.46\% $\pm$ 1.54\% \\
Qwen3-Embedding-0.6B-Q8\_0    & 300.0 &  202.1 & 781.65 & 84.26\% $\pm$ 1.47\% \\
embeddinggemma-300m-Q4\_0     & 409.2 & 1219.4 & 409.66 & 79.51\% $\pm$ 1.63\% \\
embeddinggemma-300M-Q8\_0     & 316.2 &  919.4 & 402.16 & 78.69\% $\pm$ 1.66\% \\
embeddinggemma-300M-BF16      & 361.8 &  886.0 & 672.88 & 78.36\% $\pm$ 1.67\% \\
\bottomrule
\end{tabular}
}
\end{table}
\FloatBarrier

The data demonstrates that all evaluated candidates achieve strong baseline accuracy for tool classification, with scores ranging from 78.20\% to 84.26\%. The Q8-quantized variant of \texttt{Qwen3-Embedding-0.6B} achieved the highest raw accuracy of 84.26\% ($\pm$ 1.47\%), outperforming the Jina-v5 and Gemma model families. However, this accuracy advantage is accompanied by significant efficiency trade-offs: the Qwen models exhibit the largest memory footprints (750--781 MB) and low throughput (202--251 tokens per second). Conversely, Jina's nano models demonstrate exceptional throughput, with the Q4 variant reaching 1380.0 tokens per second while maintaining an accuracy of 80.49\%. 

Because tool retrieval represents a relatively simple classification task where all models maintain a high baseline, it cannot serve as the sole selection metric. Therefore, the models must be further compared on the more complex task of option retrieval.

\subsubsection{Task 2: Option Retrieval Results}

The second task evaluated the models on their ability to retrieve the correct set of command-line flags (options) for a given query. This task was split into two stages: a baseline evaluation utilizing the raw validation queries, and an optimized grid-search evaluation sweeping all combinations of query-tool prepending, flag title formatting, and catalog keyword augmentation. For both stages, the ranked options were partitioned using a dynamic Otsu cutoff to identify the group of best matches. The baseline metric results  are compiled in Table~\ref{tab:option_retrieval_baseline}, and the optimized results (the best configuration per model) are shown in Table~\ref{tab:option_retrieval_optimized}.

\FloatBarrier
\begin{table}[htbp]
\centering
\caption{Task 2: Option Retrieval Baseline Results (No Optimizations)}
\label{tab:option_retrieval_baseline}
\resizebox{\textwidth}{!}{
\begin{tabular}{l r r r r r}
\toprule
\textbf{Model Name} & \textbf{Precision (\%)} & \textbf{Recall (\%)} & \textbf{F1 (\%)} & \textbf{RSCE} & \textbf{Recall/RSCE} \\
\midrule
v5-nano-retrieval-Q4\_K\_M    & 27.38\% & 43.47\% & 26.22\% & 3.40 & 12.79 \\
v5-nano-retrieval-Q8\_0       & 28.38\% & 45.27\% & 27.28\% & 3.22 & 14.06 \\
v5-nano-retrieval-F16         & 27.96\% & 46.01\% & 27.12\% & 3.26 & 14.11 \\
v5-small-retrieval-Q4\_K\_M   & 20.52\% & 38.72\% & 20.91\% & 3.62 & 10.70 \\
v5-small-retrieval-Q8\_0      & 20.98\% & 41.48\% & 21.93\% & 3.69 & 11.24 \\
qwen3-embedding-0.6b-q4\_k\_m & 27.34\% & 38.06\% & 26.16\% & 2.69 & 14.15 \\
Qwen3-Embedding-0.6B-Q8\_0    & 17.20\% & 46.72\% & 19.93\% & 4.36 & 10.72 \\
embeddinggemma-300m-Q4\_0     & 37.32\% & 38.01\% & 33.13\% & 2.35 & 16.17 \\
embeddinggemma-300M-Q8\_0     & 38.41\% & 35.87\% & 33.27\% & 2.12 & 16.92 \\
embeddinggemma-300M-BF16      & 38.05\% & 37.84\% & 33.58\% & 2.26 & 16.74 \\
\bottomrule
\end{tabular}
}
\end{table}
\FloatBarrier

\FloatBarrier
\begin{table}[htbp]
\centering
\caption{Task 2: Option Retrieval Optimized Results}
\label{tab:option_retrieval_optimized}
\resizebox{\textwidth}{!}{
\begin{tabular}{l r r r r r}
\toprule
\textbf{Model Name} & \textbf{Precision (\%)} & \textbf{Recall (\%)} & \textbf{F1 (\%)} & \textbf{RSCE} & \textbf{Recall/RSCE} \\
\midrule
v5-nano-retrieval-Q4\_K\_M    & 31.70\% & 57.81\% & 32.51\% & 3.73 & 15.50 \\
v5-nano-retrieval-Q8\_0       & 32.53\% & 53.85\% & 32.61\% & 3.44 & 15.65 \\
v5-nano-retrieval-F16         & 31.79\% & 52.98\% & 32.09\% & 3.38 & 15.67 \\
v5-small-retrieval-Q4\_K\_M   & 28.84\% & 45.08\% & 27.63\% & 3.40 & 13.26 \\
v5-small-retrieval-Q8\_0      & 28.19\% & 47.51\% & 27.51\% & 3.57 & 13.31 \\
qwen3-embedding-0.6b-q4\_k\_m & 49.04\% & 49.84\% & 42.27\% & 2.66 & 18.74 \\
Qwen3-Embedding-0.6B-Q8\_0    & 44.37\% & 54.95\% & 40.78\% & 2.98 & 18.44 \\
embeddinggemma-300m-Q4\_0     & 49.20\% & 48.25\% & 42.85\% & 2.39 & 20.19 \\
embeddinggemma-300M-Q8\_0     & 50.42\% & 48.47\% & 43.47\% & 2.39 & 20.28 \\
embeddinggemma-300M-BF16      & 51.77\% & 48.93\% & 44.61\% & 2.31 & 21.18 \\
\bottomrule
\end{tabular}
}
\end{table}
\FloatBarrier

The baseline evaluation reveals that Google's \texttt{embeddinggemma-300m} family performs exceptionally well out-of-the-box, with the Q8 variant achieving a baseline F1-score of 33.27\% and a Recall/RSCE score of 16.92. This baseline strength is largely due to the title-text structure inherent in Gemma's document ingestion. In contrast, Qwen and Jina struggle under baseline conditions, exhibiting F1-scores between 19.93\% and 27.28\% and higher \ac{RSCE} scores.

Upon applying the query and catalog optimizations, all models showed substantial performance improvements, validating the optimization methodology:
\begin{itemize}
    \item \textbf{Qwen-0.6B (RUN 5)}: Appending the target tool name to the query and injecting titles into option descriptions resolved semantic confusion, resulting in a dramatic F1-score increase for Qwen3-Q4 from 26.16\% to 42.27\%, and a Recall/RSCE improvement from 14.15 to 18.74. 
    \item \textbf{Jina-v5 (RUN 5/6/8)}: The Jina models achieved their optimal configurations when incorporating both query-tool appending and keyword descriptions, showing F1-score increases of approximately 5--6\%.
    \item \textbf{Gemma-300M (RUN 6/8)}: The Gemma family achieved the highest optimized Recall/RSCE scores across all candidates, with the BF16 variant reaching 21.18 (F1: 44.61\%, RSCE: 2.31) and the Q8 variant reaching 20.28 (F1: 43.47\%, RSCE: 2.39). Notably, for Gemma, RUN 6 and RUN 8 yielded identical scores. Because the option flag name is natively required to form Gemma's document title field during embedding generation, the option name optimization is active by default in both runs, rendering the explicit `--opt-desc-name` CLI flag redundant.
\end{itemize}

\subsubsection{Resource and Computational Efficiency}

Because the target command retrieval system must execute locally on resource-constrained devices, raw retrieval quality must be balanced against computational efficiency. 
The Jina-v5-nano models are the most lightweight, maintaining a peak memory footprint of approximately 280--300 MB for the quantized versions and loading in under 300 ms. However, this small memory footprint comes at the expense of retrieval quality, as Jina-nano models achieved the lowest optimized F1-scores (~32.6\%). 

At the other extreme, the \texttt{Qwen3-Embedding-0.6B} variants demand high computational resources, consuming up to 781.65 MB of peak memory (8 bit quantized) and exhibiting low throughput. While Qwen offers high precision, the execution latency and resource requirements make it less suitable for time efficient local search.

The \texttt{embeddinggemma-300m} family offers an optimal compromise. The BF16 variant achieves the highest retrieval accuracy but requires 672.88 MB of memory. Quantizing the model to 8-bit integer precision (\texttt{Q8\_0}) reduces the peak memory consumption by 37.3\% (down to 402.16 MB) and increases throughput from 886.0 to 919.4 tokens per second, while maintaining 95.7\% of the unquantized model's Recall/RSCE score (20.28 vs 21.18). 

\subsubsection{Weighted Multi-Criteria Decision Analysis}

To determine the optimal model configuration, a multi-criteria decision matrix was constructed using Min-Max utility scaling normalization. For each candidate configuration $i$ and metric $j$, a normalized utility score $u_{i,j} \in [0, 10]$ was calculated. Positive metrics (where higher values are better) were normalized according to:

\begin{equation}
u_{i,j} = 10 \times \frac{x_{i,j} - \min_k(x_{k,j})}{\max_k(x_{k,j}) - \min_k(x_{k,j})}
\end{equation}

For negative metrics (where lower values are better, such as peak memory and model load time), the utility score was defined as:

\begin{equation}
u_{i,j} = 10 \times \frac{\max_k(x_{k,j}) - x_{i,j}}{\max_k(x_{k,j}) - \min_k(x_{k,j})}
\end{equation}

The criteria weights were defined as follows: Recall/RSCE ($w=3$) to prioritize retrieval balance; throughput ($w=2$), peak memory ($w=2$), and load time ($w=3$) to emphasize computational efficiency; and F1-score ($w=1$), precision ($w=1$), recall ($w=1$), and tool accuracy ($w=1$) as baseline quality indicators. The model load time was assigned a weight of 3 because in local systems executing short natural-language queries, startup latency is a dominant factor in end-to-end user experience. The normalized scores and final weighted sums are presented in Table~\ref{tab:decision_matrix}.

\FloatBarrier
\begin{table}[htbp]
\centering
\caption{Model Selection Decision Matrix (Min-Max Utility Scaling)}
\label{tab:decision_matrix}
\resizebox{\textwidth}{!}{
\begin{tabular}{l *{8}{r} | r r r}
\toprule
\textbf{Model Variant} & 
\textbf{Recall/RSCE} & 
\textbf{Throughput (TPS)} & 
\textbf{Load Time (ms)} & 
\textbf{Peak Memory (MB)} & 
\textbf{F1-Score} & 
\textbf{Precision} & 
\textbf{Recall} & 
\textbf{Tool Accuracy} & 
\textbf{Weighted Sum} & 
\textbf{Final Score} & 
\textbf{Rank} \\
\midrule
Gemma-300M-Q8       & 8.86 &  6.13 &  9.61 & 7.60 &  9.33 &  9.43 & 2.66 & 0.81 & 105.10 & 7.51 & \textbf{1} \\
Gemma-300m-Q4       & 8.75 &  8.65 &  7.48 & 7.45 &  8.97 &  8.91 & 2.49 & 2.16 & 103.41 & 7.39 & 2 \\
Gemma-300M-BF16     & 10.00 &  5.85 &  8.57 & 2.18 & 10.00 & 10.00 & 3.02 & 0.26 &  95.04 & 6.79 & 3 \\
v5-nano-Q4\_K\_M    & 2.83 & 10.00 &  7.79 & 10.00 & 2.92 &  1.49 & 10.00 & 3.78 &  90.05 & 6.43 & 4 \\
v5-nano-Q8\_0       & 3.02 &  5.93 & 10.00 & 9.58 &  2.98 &  1.84 & 6.89 & 3.78 &  85.59 & 6.11 & 5 \\
Qwen3-Q8            & 6.54 &  0.11 &  9.98 & 0.00 &  7.76 &  6.86 & 7.75 & 10.00 &  82.14 & 5.87 & 6 \\
v5-nano-F16         & 3.05 &  4.64 &  9.89 & 5.74 &  2.68 &  1.53 & 6.21 & 4.06 &  74.04 & 5.29 & 7 \\
qwen3-q4            & 6.91 &  0.52 &  3.48 & 0.57 &  8.63 &  8.84 & 3.74 & 7.03 &  61.58 & 4.40 & 8 \\
v5-small-Q8         & 0.06 &  0.00 &  9.86 & 0.27 &  0.00 &  0.00 & 1.91 & 1.07 &  33.28 & 2.38 & 9 \\
v5-small-Q4         & 0.00 &  0.50 &  0.00 & 1.03 &  0.07 &  0.28 & 0.00 & 0.00 &   3.40 & 0.24 & 10 \\
\bottomrule
\end{tabular}
}
\end{table}
\FloatBarrier

The decision matrix identifies \texttt{embeddinggemma-300M-Q8\_0} as the optimal model configuration, achieving the highest final weighted sum 105.10. It is followed closely by the Q4-quantized variant of the same base model (Weighted Sum: 103.42). This result confirms that the Q8-quantized Gemma model successfully balances retrieval quality with runtime speed and low resource usage.

To visually illustrate the trade-offs involved in this selection, Figure~\ref{fig:pareto_frontier} plots retrieval quality (Recall/RSCE) against peak memory consumption. The dashed red line represents the Pareto-optimal frontier, indicating configurations that cannot be improved in one metric without degrading the other.
\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    title={Pareto Frontier: Retrieval Quality vs. Memory Consumption},
    xlabel={Peak Memory Consumption (MB)},
    ylabel={Retrieval Quality Score (Recall/RSCE)},
    xmin=200, xmax=900,
    ymin=10, ymax=23,
    xtick={300, 400, 500, 600, 700, 800},
    ytick={12, 14, 16, 18, 20, 22},
    legend pos=south west,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt, draw=gray!20},
    width=0.9\textwidth,
    height=7.5cm
]

% Pareto frontier line connecting optimal points (Jina Q4 -> Gemma Q8 -> Gemma BF16)
\addplot[
    color=primaryred,
    dashed,
    thick,
    mark=none
] coordinates {
    (282.1, 15.5)
    (402.1, 20.28)
    (672.8, 21.18)
};
\addlegendentry{Pareto Frontier}

% Jina Models (Blue)
\addplot[
    only marks,
    color=primaryblue,
    mark=square*,
    mark size=2.5pt
] coordinates {
    (282.1, 15.50) % v5-nano-Q4
    (302.9, 15.65) % v5-nano-Q8
    (494.8, 15.67) % v5-nano-F16
    (730.2, 13.26) % v5-small-Q4
    (768.4, 13.31) % v5-small-Q8
};
\addlegendentry{Jina-v5 Variants}

% Qwen Models (Orange)
\addplot[
    only marks,
    color=primaryorange,
    mark=triangle*,
    mark size=3pt
] coordinates {
    (753.3, 18.74) % Qwen3-Q4
    (781.6, 18.44) % Qwen3-Q8
};
\addlegendentry{Qwen3 Variants}

% Gemma Models (Green)
\addplot[
    only marks,
    color=primarygreen,
    mark=*,
    mark size=2.5pt
] coordinates {
    (409.6, 20.19) % Gemma-Q4
    (402.1, 20.28) % Gemma-Q8
    (672.8, 21.18) % Gemma-BF16
};
\addlegendentry{Gemma-300M Variants}

% Labels for key models (Adjusted to stay within plot bounds and clear the legend)
\node[above, font=\scriptsize, yshift=4pt] at (axis cs:282.1, 15.50) {Jina Nano Q4};
\node[right, font=\scriptsize, xshift=4pt, yshift=2pt] at (axis cs:302.9, 15.65) {Jina Nano Q8};
\node[right, font=\scriptsize, xshift=4pt] at (axis cs:781.6, 18.44) {Qwen3 Q8};
\node[above, font=\scriptsize, yshift=3pt] at (axis cs:402.1, 20.28) {Gemma Q8 (Winner)};
\node[above, font=\scriptsize, yshift=3pt] at (axis cs:672.8, 21.18) {Gemma BF16};

\end{axis}
\end{tikzpicture}
\caption{Retrieval quality (Recall/RSCE) plotted against peak memory consumption (VmHWM in MB). The Pareto frontier (dashed line) shows the optimal configurations, highlighting that the Q8-quantized Gemma-300M model represents the Pareto-optimal design point under strict memory and latency constraints.}
\label{fig:pareto_frontier}
\end{figure}
The Pareto chart clearly shows that while the Jina-v5-nano models minimize memory consumption and Gemma-300M-BF16 maximizes retrieval quality, \texttt{embeddinggemma-300M-Q8\_0} occupies a critical elbow position. It offers near-peak retrieval quality at a resource cost comparable to the much smaller Jina models. Based on this quantitative analysis and trade-off visualization, \textbf{\texttt{embeddinggemma-300M-Q8\_0} is selected as the primary embedding engine} for the command retrieval system.
