\section{Conclusion And Future Work}

The thesis successfully answers the research questions. A functional PoC system was designed that operates within the required constraints. Whilst only retrieving the accurate options in a fraction of absolute cases it retrieved a subset of the required options much more frequently and successfully failed command generation if conflicting options were selected. This lead to only about 21.4\% of cases where no correct options were suggested at all with only 9.2\% confidently suggesting a wrong tool without causing errors. This shows the potential of the system especially as a tool for learning if suggested options always were suggested with a fitting explaination to allow the user to verify the suggested command actually does what they want. So while this tool will not help at solving the problem of poor documentation it can aid in assisting the understanding and simple utilization of comand line tools without requiring users to read man pages.

Future works should evaluate the performance of Specialized neural networks as a substitutional engine and maybe also the performance of cross encoders as a reranker for the ranked options, using a cross encoder as a direct retrieval engine might lead to longer response times and is not recommended only alternative would be to build the cross encoder directly into the other retrieval engines but this might lead to increased load times and should only done with careful consideration. Furthermore experiments to build a Classifier model that detects false matches might significantly increase performance. It is recommended to keep improving this approach by adding and determining the missing weights to the individual engines. Finally a more broad user study with a representative statistically reliable cohort should be conducted. 
Furthermore soem of the results of this work are directly transferable to the implementation and research of creating localized RAG pipelines with strict performance boundries since essentially the proposed system implements RAG just without utilizing a LLM for the final command generation.

% ====================================================================
% Draft of the Conclusion & Future Work Sections
% ====================================================================

\subsection{Conclusion}

This work demonstrated the viability of a local-first, privacy-preserving command option retrieval system operating under strict offline and hardware constraints. Technical benchmarking showed that the classification model routed queries to the correct tool in 80.30\% of cases, leading to a successful command resolution rate of 58.13\%. In terms of flag retrieval, the system returned a perfect match in 24.16\% of successful resolutions and a partial match in 51.01\% of cases. Crucially, the system proved mostly safe: only 21.35\% of all test cases resulted in a generated command that offered no correct options, and only 9.20\% generated incorrect commands suggesting a wrong tool confidently without raising syntax errors. The remaining failures were safely intercepted by the systems validation engine, which prevented the suggestion of conflicting options. From a performance standpoint, the system achieved a sub-second average latency of 563.21~ms and maintained a stable memory footprint of 658.62~MB on standard CPU hardware. The subsequent user study confirmed that participants valued the reduction in cognitive load and manual searching effort, though varying confidence in the retrieved recommendations underscores the need for transparency in AI-assisted workflows.

\subsection{Practical Recommendations}

Based on these findings, several recommendations are proposed for organizations and developers implementing local-first retrieval systems. First, to mitigate the manual labor required to construct and update curated tool catalogs, developers should integrate semi-automated extraction pipelines that parse option definitions directly from source code, abstract syntax trees, or standard documentation. Second, to foster trust and ensure execution safety, the user interface must never present generated command strings without an accompanying natural language description of the tool and options suggested. This allows for manual verification and helps users to trust the system. Third, combining semantic search with strict local syntax validation is essential. The system should fail explicitly and informatively rather than proposing syntactically invalid commands.

\subsection{Future Work}

Future research should focus on refining the retrieval architecture and validating its generalizability. To improve matching accuracy, specialized neural networks should be evaluated as supplementary engines to the existing retrieval framework. Additionally, cross-encoders could be explored as secondary rerankers to optimize option ranking, though this requires careful consideration to avoid excessive loading times or latency. A dedicated binary classifier model could also be developed to detect false matches and intercept incorrect routing decisions before command generation. Furthermore, the collaborative retrieval system can be optimized by mathematically calibrating the individual weights of the retrieval engines instead of relying on uniform score fusion. To obtain generalizable usability metrics, future evaluations must conduct user studies with a larger, statistically representative cohort and expand the benchmark dataset to cover a wider selection of command-line utilities. Finally, the design patterns and multi-engine aggregation developed in this work are directly transferable to the implementation of resource-constrained, localized Retrieval-Augmented Generation systems.
