
\section{Implementation}

The implementation of the tool was based on the system design and technology decisions of cahpter~\ref{sec:design_and_decision}. We utilized the tool Antigravity CLI in combination with various Gemini \acp{LLM} to accelerate development. For this we first created a project scaffold, with placholder functionality. This allowed us to ensure the architecture outlined in the respective chapters was properly implemented. We then utilized Antigravity CLI to implement the actual functionality of the tool step by step. By keeping tight boundries on what should be created and manually verifying the correctness of generated code, we managed to build a functional tool, fulfilling a subset of the requirements specified. In subsection \ref{sec:implementation_limitation} we outline the requirements ommitted at this stage and the reasons behind it.

\subsection{Score Aggregation}\label{sec:score_aggregation_impl}

To consolidate candidates retrieved by multiple matching engines, the system normalizes and aggregates raw scores using weighted summation. First, raw scores $S_e(c)$ for candidate $c$ from engine $e$ are normalized against the maximum score returned by that specific engine to bound them to $[0, 1]$:
\[
S'_{e}(c) = \begin{cases} 
\frac{S_e(c)}{\max_{c'} S_e(c')} & \text{if } \max_{c'} S_e(c') > 0 \\ 
0.0 & \text{otherwise} 
\end{cases}
\]
Next, the final aggregated score $S_{\text{aggregated}}(c)$ is computed as the weighted sum:
\[
S_{\text{aggregated}}(c) = \sum_{e \in E} S'_{e}(c) \cdot w_e
\]
where $w_e \ge 0$ is the pre-configured weight of engine $e$. Candidates missing from an engine's results receive a score of $0.0$ for that engine. Finally, candidates are consolidated and sorted in descending order of their aggregated scores to produce the final ranked list.

\subsection{Hyperparameter Optimization for Otsu-based Candidate Cutoff}
\label{sec:otsu_optimization}

To select the best matching candidate group, candidate options are pruned using an Otsu-based cutoff filter. As detailed in Section~\ref{sec:score_aggregation_impl}, similarity scores from the keyword and embedding matching engines are normalized via min-max scaling and combined using a weighted sum, to which the cutoff filter is applied directly. 

The filter is parameterized by the relative floor $\alpha \in [0, 1]$ (which zeroes out candidates scoring below $\alpha \cdot S_{\text{max}}$) and the Otsu multiplier $m \ge 1.0$ (which scales the computed Otsu drop threshold $\theta = m \cdot \theta_{\text{Otsu}}$ to determine where the score ``cliff'' occurs).

A grid search was conducted over $\alpha \in [0.5, 1.5]$ (step 0.1) and $m \in [1.0, 3.0]$ (step 0.5) across $N=610$ queries. Performance was measured using Precision, Recall, $F_1$-Score, and \ac{RSCE} (defined in Section~\ref{sec:model_selection_meth}) alongside the Recall/\ac{RSCE} ratio. Table~\ref{tab:otsu_grid_search} summarizes the top configurations.

\begin{table}[htbp]
\centering
\caption{Otsu cutoff grid search top configurations.}
\label{tab:otsu_grid_search}
\begin{tabular}{ccccccc}
\toprule
$\alpha$ & $m$ & Avg. Precision & Avg. Recall & Avg. $F_1$-Score & Avg. RSCE & Avg. Recall/RSCE \\
\midrule
\multicolumn{7}{l}{\textbf{Recall-Optimized}} \\
0.50 & 3.00 & 0.3114 & 0.4833 & 0.3378 & 2.2590 & 0.2572 \\
0.50 & 2.50 & 0.3121 & 0.4787 & 0.3375 & 2.2301 & 0.2580 \\
0.50 & 2.00 & 0.3156 & 0.4691 & 0.3383 & 2.1481 & 0.2610 \\
\midrule
\multicolumn{7}{l}{\textbf{RSCE-Optimized}} \\
0.90 & 3.00 & 0.3657 & 0.3077 & 0.3143 & 1.6322 & 0.2267 \\
0.90 & 2.00 & 0.3649 & 0.3068 & 0.3135 & 1.6347 & 0.2259 \\
0.90 & 1.00 & 0.3655 & 0.3077 & 0.3141 & 1.6363 & 0.2265 \\
\midrule
\multicolumn{7}{l}{\textbf{Recall/RSCE-Optimized}} \\
0.60 & 2.00 & 0.3394 & 0.3954 & 0.3381 & 1.7470 & 0.2708 \\
0.60 & 1.50 & 0.3391 & 0.3945 & 0.3376 & 1.7454 & 0.2705 \\
0.60 & 3.00 & 0.3378 & 0.3967 & 0.3374 & 1.7598 & 0.2697 \\
\bottomrule
\end{tabular}
\end{table}

We selected the configuration $\alpha = 0.50, m = 3.00$ (Recall-optimized). Optimizing for RSCE ($\alpha = 0.90$) or Recall/RSCE ($\alpha = 0.60$) severely degrades option recall (dropping from $0.4833$ to $0.3077$ and $0.3954$, respectively), which is unacceptable for identifying critical options. The higher count error associated with the selected configuration (\ac{RSCE} = 2.2590) is partially mitigated downstream by the Syntactical Validator (Section~\ref{sec:impl_validator}), which filters out invalid, conflicting, or non-applicable option combinations.


\subsection{Transition-Driven Command Assembly and Validation}\label{sec:impl_validator}

To ensure that generated CLI parameters compile into syntactically valid and executable statements, we implement a \ac{STRS} rather than a passive parser. Traditional parser generators act as passive recognizers, accepting or rejecting complete, pre-ordered token sequences. In contrast, the \ac{STRS} acts as an active, transition-driven command assembler. It accepts an unordered, raw, or incomplete set of generated options, dynamically restructures the token queue, auto-completes missing defaults, resolves mutual exclusions, and compiles the final elements into a command string. The declarative syntax and validation boundaries of the STRS are defined entirely within the \texttt{rules} field of the catalog files (see Appendix **appendix** for the schemas of \texttt{cp-catalog.json}, \texttt{mv-catalog.json}, and \texttt{ops-sync-catalog.json}). 

The grammar of the STRS is structured as a transition table mapping active states to target states based on token matches. The process begins in an initial state (typically \texttt{START}) and must terminate in an accepting state (e.g. \texttt{END\_STATE}). Transition matches can be exact literals or regular expressions (\texttt{isRegex}), which allows the system to validate dynamic inputs. To prevent conflicting flags, a transition can define an exclusion list (\texttt{excludes}) containing option IDs. If any option listed in this exclusion list has already been processed during the execution, the validation fails immediately, preventing invalid flag combinations.

Context-sensitive reordering is resolved via two dynamic queue-manipulation operations: token pulling and token injection. Token pulling (\texttt{pull}) addresses order-dependency requirements where the occurrence of a specific flag shifts the expected location of another token. When a transition triggers a pull, the rewriter scans the remaining input queue, extracts the target token from its current index, and pulls it to the head of the queue to satisfy the immediate transition constraint. Conversely, token injection (\texttt{injection}) inserts implicit options or default arguments at either the front of the queue (\texttt{Immediate}) or the back (\texttt{End}). The rewriter supports both whitespace separation and token fusion via custom delimiters, especially relevant for injecting argument values correctly. To prevent infinite loops or parameter redundancy, candidate injections are validated against a registry of currently queued tokens, processed raw tokens, and triggered option IDs, skipping any duplicate entries. Once the input queue is exhausted, the \ac{STRS} accepts the command if the current state is an accepting state, outputting the buffer as a space-separated string representing the assembled command.



\subsection{Compilation}

To ensure portability, performance and avoid compilation issues, we did not use the standard \texttt{cargo build --release} command for compilation. We instead used the following:

\texttt{CFLAGS="-Du\_int8\_t=uint8\_t -Du\_int16\_t=uint16\_t -Du\_int64\_t=uint64\_t" RUSTFLAGS="-C target-feature=+avx,+avx2,+fma,+f16c,+bmi2" cargo zigbuild --release --target x86\_64-unknown-linux-musl}

The command can be split into a total of five sections:

\begin{enumerate}
    \item \textbf{CFLAGS="-Du\_int8\_t=uint8\_t -Du\_int16\_t=uint16\_t -Du\_int64\_t=uint64\_t"} 
    
    Because the used version of sqlite-vec (v0.1.9) utilizes legacy types (e.g. u\_int8\_t) **source** this part of the compilation command defines a preprocessor macro for the compiler to map those legacy types to C99 standard types (e.g. uint8\_t). This is required because otherwise the musl build fails.
    \item \textbf{RUSTFLAGS="-C target-feature=+avx,+avx2,+fma,+f16c,+bmi2"}
    
    This part of the command enables hardware-level CPU instruction extensions on the target machine. The \textbf{avx} and \textbf{avx2} extensions allow the CPU to run \ac{SIMD} operations, processing up to 8 floating-point numbers in a single clock cycle instead of one by one **source**. On the other hand the \textbf{fma} allows teh CPU to perform multiplication and addition (A x B + C) in a single CPU instruction. This is the fundamental math operation behind neural networks and matrix multiplications **source**, leading to a significant speedup of those operations. The \textbf{f16c} extension enables hardware acceleration for converting 16-bit floats to 32-bit floats, which may help to speed up decompression of machine learning models **source**. Finally the \textbf{bmi2} extension accelerates low-level indexing and hashing operations **source**.
    \item \textbf{cargo zigbuild}
    
    By using this part in our comamnd we swap out cargos default linker and C compiler with Zig. This allows for teh cross compilation of C and C++ code without requiring a complex setup of a copilation environment **source**. 
    \item \textbf{--release}
    
    Using this flag in our compilation command, tells cargo to optimize code, strip debugging symbols and trun of runtime debugging assertions, leading to an optimized executable.
    \item \textbf{--target x86\_64-unknown-linux-musl}
    
    This last part of the command, specifies the compilation target system. This specific variation will compile for an x86 64 bit system. By specifying linux it is specified that the executable targets the a Linux kernel based OS and using musl tells the compiler to use musl-libc instead of the standard glibc. Unlike the standard gnu target, the  musl  target configures the compiler to link the C standard library statically **source**, making our binary portable and not require any shared runtime dependencies.
\end{enumerate}

% - Rust

% Beschreibung der Module und wie sie ineinandergreifen (Diagram?)

% Ggf. eine beschreibung wie man besonders effectiv die Beschreibungen für das Embedding formuliert.

% tar, mv, cp, 

\subsection{Limitations}\label{sec:implementation_limitation}

Due to time constraint a selection of non-essential requirements were ommitted during the implementation phase. In particular, no authentication (FR-3) and access control mechanisms (NFR-11) were implemented at this time. Neither was the concurrency resilience (NFR-12). This decision was made, because the application is not supposed to be deployed in a production environment at this stage, hence making security features a secondary concern. Furthermore, the interactive disambiguatio (FR-7) was ommitted for now. Users are still notified when a conflict has been detected, but the program does not enter another execution cycle after that. Instead it finishes it's run, which requires the engines to be reloaded, if a user revises their query and calls the program an additional time. We decided that this was a negligible concern, since it does not directly influence the usefulness of the program. Finally the configuration management (FR-9) and diagnostic logging were not implemented at this stage, thereby also causing NFR-14 to become redundant. We decided to omit those features, because at this stage they do not aid the usability of the program and no diagnostics are required.