\section{Systems Design And Implementation}

\subsection{Design}

Database:
- Tabelle für Tool beschreibungen
    - Tool name (ID)
    - Tool Beschreibung (TEXT)
    - Tool Beschreibung Vector (BLOB)
- Tabelle für tool option Beschreibungen
    - option (ID)
    - Tool Name (ID)
    - Tool Option beschreibung (TEXT)
    - Tool Option Beschreibung Vector (BLOB)

Pipeline Design \& Systemablaufdiagramm

2 systeme, eine lightweight api die das modell lädt und embeddings macht falls app registriert wird und das cli system



\subsubsection{Requirements}

\begin{table}[htbp]
    \centering
    \caption{Hardware and Environmental Constraints}
    \label{tab:env_constraints}
    \renewcommand{\arraystretch}{1.3} % Adds slightly more space between rows
    \begin{tabularx}{\textwidth}{@{} l l X @{}}
        \toprule
        \textbf{ID} & \textbf{Constraint} & \textbf{Description} \\
        \midrule
        ENV-1 & Target Hardware & All performance metrics must be achieved on a baseline Proxmox VM configured with 4 vCPUs (AMD Ryzen 7 5700X architecture), 8GB of 2133MHz DDR4 RAM, and a Samsung SSD 970 Evo Plus NVMe drive. \\
        ENV-2 & Offline Isolation & The system must function completely offline. It must not initiate any outbound network requests under any circumstances. \\
        ENV-3 & Zero Third-Party Daemon & The system must be entirely self-contained. While the system's own architecture may utilize background processes, it must not require the installation or execution of third-party standalone server software. \\
        \bottomrule
    \end{tabularx}
\end{table}

\FloatBarrier

\begin{table}[htbp]
    \centering
    \caption{Functional Requirements (FR)}
    \label{tab:functional_reqs}
    \renewcommand{\arraystretch}{1.3}
    \begin{tabularx}{\textwidth}{@{} l l X @{}}
        \toprule
        \textbf{ID} & \textbf{Requirement} & \textbf{Description} \\
        \midrule
        FR-1 & Command Generation & The system must accept natural language queries and return a syntactically correct command-line string comprising the base tool, relevant option flags, and placeholders for user-specific positional parameters. \\
        FR-2 & Non-Execution Boundary & The system must strictly output the generated command as text and never autonomously execute it in the host shell. \\
        FR-3 & Authenticated Ingestion & The system must provide a programmatic interface to ingest and index new tool catalogs, utilizing an authentication or cryptographic authorization mechanism to prevent unauthorized manipulation. \\
        FR-4 & Catalog Lifecycle Management & The system must provide authenticated interfaces allowing authorized entities to list, update, and remove previously ingested tool catalogs. \\
        FR-5 & State Persistence & The system must persist all ingested catalogs and computed retrieval states locally, ensuring immediate availability across system reboots and terminal sessions without requiring re-ingestion. \\
        FR-6 & Syntactical Validation & The system must logically evaluate generated commands to ensure no mutually exclusive flags are recommended simultaneously, and that any required positional ordering is strictly respected. \\
        FR-7 & Interactive Disambiguation & If a query is ambiguous, triggers a flag conflict, or falls outside the domain of ingested catalogs, the system must halt and prompt the user for clarification rather than returning an invalid command. \\
        \bottomrule
    \end{tabularx}
\end{table}

\FloatBarrier
\renewcommand{\arraystretch}{1.3} % You can safely increase this again for readability
\begin{xltabular}{\textwidth}{@{} l >{\raggedright\arraybackslash}p{2.8cm} >{\raggedright\arraybackslash}X @{}}

    % --- 1. FIRST PAGE HEADER ---
    \caption{Non-Functional Requirements (NFR)}
    \label{tab:non_functional_reqs} \\
    \toprule
    \textbf{ID} & \textbf{Requirement} & \textbf{Description} \\
    \midrule
    \endfirsthead

    % --- 2. SUBSEQUENT PAGE HEADERS ---
    \caption*{Table \ref{tab:non_functional_reqs}: Non-Functional Requirements (Continued)} \\
    \toprule
    \textbf{ID} & \textbf{Requirement} & \textbf{Description} \\
    \midrule
    \endhead

    % --- 3. BOTTOM OF PAGE FOOTER (When breaking to a new page) ---
    \midrule
    \multicolumn{3}{r}{\textit{Continued on next page \dots}} \\
    \endfoot

    % --- 4. ABSOLUTE LAST FOOTER (End of the entire table) ---
    \bottomrule
    \endlastfoot

    % --- DATA ROWS ---
    NFR-1 & Query Latency & The system must return a completed command recommendation in under 1000ms from the moment the query is submitted. \\
    
    NFR-2 & Cold Startup Latency & The application must initialize and be ready to process input in under $1000ms$. \\
    
    NFR-3 & Memory Bounding & The peak RAM utilization of the system process must never exceed $1024 MB$ at any time during operation. \\
    
    NFR-4 & Latency Scalability & The system must maintain its $< 1000ms$ query latency when the persistent state is scaled to hold definitions for up to ** Average number of CLI tools installed on a linux machine ** distinct tools. \\
    
    NFR-5 & Retrieval Quality & The system must achieve a minimum Macro-Averaged F1-Score of **[Target]** and a Jaccard Similarity of **[Target]** against the established evaluation dataset. \\
    
    NFR-6 & Storage Modularity & The system architecture must strictly decouple the core application logic from the persistent state mechanisms. It must be possible to completely replace the underlying storage backend without modifying the domain logic, retrieval algorithms, or user interfaces. \\

    NFR-7 & Search and Matching Modularity & The system architecture must isolate all specific search and matching mechanisms behind abstract interfaces.\\
    
    NFR-8 & Similarity and Rank Aggregation & The system must support the multi-strategy evaluation of candidate command-line options against the user's natural language query. It must provide an internal mechanism to mathematically normalize and fuse disparate similarity scores from independent matching algorithms into a single, unified ranking before outputting the finalized result.\\

    NFR-9 & Zero-Runtime Portability & The system must be deployable as self-contained artifacts for Linux, macOS, and Windows ** MAYBE LIMIT TO LINUX **, requiring no pre-installed language runtimes on the host machine. \\
    
    NFR-10 & IPC Network Isolation & Any \ac{IPC} mechanisms utilized by the system must be strictly bound to local OS constructs, explicitly denying any non-local network access. \\
    
    NFR-11 & State Access Control & The system must utilize host OS access control mechanisms to ensure its persistent state data is logically isolated and protected from unauthorized modification by other non-privileged users. \\
    
    NFR-12 & Concurrency Resilience & The system must safely process simultaneous execution requests from multiple terminal instances without causing deadlocks, process crashes, or state corruption. \\
    
    NFR-13 & Input Sanitization & The system must sanitize all user inputs and programmatic payloads to prevent injection attacks against the underlying persistent state mechanism. \\
    
    NFR-14 & Opt-In Diagnostic Logging & By default, any diagnostic logs generated by the system must be strictly scrubbed of the user's raw natural language queries. Unredacted query logging must require explicit user opt-in. \\

\end{xltabular}

\FloatBarrier

%\subsubsection{Architectural Pattern}
%\label{sec:architecture_philosophy}

%The design of the system is fundamentally driven by the non-functional requirements governing latency, memory footprint, and modularity. Specifically, the system must process natural language queries in under 1000ms (NFR-1, NFR-4) within a 1024 MB memory limit (NFR-3), 

%**while ensuring that the core algorithmic retrieval logic remains strictly decoupled from volatile \ac{I/O} operations and storage mechanisms (NFR-6). ** Umschreiben, ggf. Requirement dass storage backend austauschbar sein muss

%To satisfy these constraints, a traditional Layered (N-Tier) Architecture was discarded. While a layered approach neatly categorizes concerns into presentation, business, and data tiers, it introduces a transitive top-down dependency structure**SOURCE**. In such a model, the core algorithmic logic is inherently coupled to the specific implementation of the underlying data access layer. 

%If the local storage mechanism were to be replaced, the business logic would require modification, directly violating the modularity constraint (NFR-6). ** UMSCHREIBEN, NFR-6 bezieht sich auf die retrieval modells (Embedding \& Bm25), ggf. NFR-6 umschreiben dass das auch darin konkret reflectiert ist! Ich möchte den Keyword matching und das embedding modell austauschen können. Ich möchte auch das storage backend tasuchen können, also ggf. 2 Requirements draus machen.

%To decouple the algorithmic core from the storage layer, the system relies on the Dependency Inversion Principle. While modern dependency-inverting paradigms such as Clean Architecture \cite{martin2012clean}**SOURCES** and Onion Architecture \cite{palermo2008onion}**SOURCES** offer robust domain isolation, they prescribe concentric internal layers that mandate extensive \ac{DTO} mapping and rigid interface boundaries. For a localized \ac{CLI} operating under strict computational and latency budgets (NFR-1, NFR-3), this enterprise-level boilerplate introduces unjustifiable computational overhead and complexity. ** Ergänzen dass die komplexität der anwendung (für die instandhaltung/erweiterung) dadurch inhärent steigt

%Therefore, this system adopts a \textbf{Hexagonal Architecture}, also known as the Ports and Adapters pattern \cite{cockburn2005hexagonal}**SOURCES**. This pattern provides the necessary macro-level boundary control without dictating internal bureaucratic layers. ** Allowing to focus on the actual implementation ** The architecture is divided into two strict conceptual zones:

%\begin{itemize}
%    \item \textbf{The Core (Inside):} This encapsulates the mathematical retrieval models, vector search logic, and domain entities (e.g., Commands, Flags, and Catalogs). The Core possesses zero knowledge of the external world. It dictates its storage and interaction needs purely through abstract interfaces (Ports).
%    \item \textbf{The Infrastructure (Outside):} This encompasses all external I/O mechanisms, including the user-facing CLI terminal (Driving Adapter) and the local persistent state database (Driven Adapter). 
%\end{itemize} 
%** Das passt so nicht ganz, Das modell und der keyword matching algorithmus müsse nauch austasuchbar sein! Daher müssen diese auch als "plugins" definiert werden.

%By implementing data-access and presentation interfaces at the edge of the Core domain,
%** Include the Embedding Modell and the Keyword matching algorithm here!**
%the external mechanisms are forced to depend on the Core rather than the inverse. This guarantees that the core natural language processing and hybrid retrieval engine can be compiled, tested, and executed entirely independent of the chosen database technology or host operating system terminal, successfully fulfilling all architectural constraints.
%** Rewrite this entire paragraph!**

\subsubsection{Architectural Pattern}
\label{sec:architecture_philosophy}

The design of the system is fundamentally driven by the non-functional requirements governing latency, memory footprint, and modularity. Specifically, the system must process natural language queries in under 1000ms (NFR-1, NFR-4) within a 1024 MB memory limit (NFR-3), while ensuring that the core domain logic remains strictly decoupled from volatile \ac{I/O} operations, allowing both the persistent storage backend (NFR-6) and the specific search and matching algorithms (NFR-7) to be seamlessly replaced.

To satisfy these constraints, a traditional Layered (N-Tier) Architecture was discarded. While a layered approach neatly categorizes concerns into different tiers (e.g. presentation, business and data), it introduces a transitive top-down dependency structure \cite{richards2020fundamentals}. As Richards and Ford assert, layers in this architecture remain isolated only if the contracts between them remain unchanged. However, in a standard layered model, the data access layer and external libraries define these contracts. Consequently, the core business logic is transitively coupled to the paradigms of the infrastructure. If the local storage technology or the specific text-matching algorithms were fundamentally replaced, the cascading changes to the underlying contracts would inevitably force direct modifications within the business logic. This top-down coupling directly violates the system's modularity constraints (NFR-6, NFR-7).

To decouple the algorithmic core from the storage and execution components, the system relies on the Dependency Inversion Principle. While modern dependency-inverting paradigms such as Clean Architecture \cite{martin2017clean} and Onion Architecture \cite{palermo2008onion} offer robust domain isolation, they prescribe concentric internal layers that mandate extensive \ac{DTO} mapping and rigid interface boundaries. For a localized command-line utility, operating under strict computational and latency budgets (NFR-1, NFR-3), this enterprise-level boilerplate introduces unjustifiable computational overhead. Furthermore, it inherently increases the structural complexity of the codebase, making future maintenance and the extension of new features unnecessarily burdensome for a utility application.

Therefore, this system adopts a \textbf{Hexagonal Architecture}, also known as the Ports and Adapters pattern \cite{cockburn2005hexagonal}. This pattern provides the necessary macro-level boundary control without dictating internal bureaucratic layers, allowing developers to maintain focus on the actual implementation of the core feature set. The architecture is divided into two strict conceptual zones:

\begin{itemize}
    \item \textbf{The Core (Inside):} This encapsulates the domain entities, the command validation (FR-6), and the mathematical normalization and score aggregation logic (NFR-8). The Core possesses zero knowledge of the external world, it dictates its storage and matching needs purely through abstract interfaces (Ports).
    \item \textbf{The Infrastructure (Outside):} This encompasses all external adapters that fulfill the Core's interfaces. This includes the user-interface, the storage backend, as well as the specific execution engines for the command option evaluation.
\end{itemize}

By implementing interface ports at the absolute boundary of the Core, the external mechanisms -- whether they are the user interface, storage backend, or the command option evaluation mechanisms -- are forced to depend on the Core rather than the inverse. This guarantees that the core natural language orchestration and rank aggregation math can be compiled, tested, and executed entirely independently of the chosen storage backend or matching technologies, successfully fulfilling all modularity constraints.


Requirement that the user must be able to interact with the system to use it and configure it.

\subsection{Implementation}

- Rust

Beschreibung der Module und wie sie ineinandergreifen (Diagram?)

Ggf. eine beschreibung wie man besonders effectiv die Beschreibungen für das Embedding formuliert.

tar, mv, cp, 