\section{Literature Review}
The following sections examine the most relevant natural-language-to-Bash systems. The reviewed work shows that command generation is possible, but also identifies limitations related to ambiguity, training data, correctness, extensibility, and resource usage. These limitations motivate the more restricted retrieval-based approach investigated in this thesis.

\subsection{Source Selection}

A publication was considered suitable when it addressed at least one central aspect of the
project. This included translating natural-language descriptions into Bash commands, retrieving technical information using natural-language queries, representing command-line syntax, or evaluating the correctness of generated commands.

Particular attention was given to the following publications:

\begin{itemize}
    \item \textit{NL2Bash: A Corpus and Semantic Parser for Natural Language Interface to the Linux Operating System} was selected because it defines one of the central datasets and baseline approaches for translating natural-language requests into Bash commands. It also discusses the difficulty of generating complete commands, including tools, options, and user-specific arguments \cite{lin2018nl2bash}.

    \item \textit{NeurIPS 2020 NLC2CMD Competition: Translating Natural Language to Bash Commands} was included because it provides a shared benchmark for the task and compares different approaches under a common evaluation setup. It is also relevant because several later publications build on the same dataset and competition results \cite{agarwal2021nlc2cmd}. 
    
    \item \textit{NL2CMD: An Updated Workflow for Natural Language to Bash Commands Translation} was selected because it extends earlier work through a revised dataset-generation process, argument masking, and Transformer-based command translation. It is particularly relevant for understanding the limitations of training-data-based approaches and the handling of command arguments \cite{fu2023nl2cmd}. 
    
    \item \textit{Explainable Natural Language to Bash Translation Using Abstract Syntax Tree} was included because it combines natural-language command generation with structural command representations and information from manual pages. This makes it relevant to the use of command documentation and to the distinction between tools, options, and arguments \cite{bharadwaj2021explainable}. 
    
    \item \textit{LLM-Supported Natural Language to Bash Translation} was selected because it examines the use of large language models for command generation and discusses weaknesses in existing datasets and evaluation methods. It is also relevant to the comparison between unrestricted command generation and the more constrained retrieval approach used in this thesis \cite{westenfelder2025llm}. 
    
    \item \textit{BashExplainer: Retrieval-Augmented Bash Code Comment Generation Based on Fine-Tuned CodeBERT} was included as adjacent work because it combines semantic and lexical retrieval for Bash-related content. Although its task direction differs from this project, it demonstrates how retrieval methods can be applied to Bash commands and how different similarity measures can complement each other \cite{yu2022bashexplainer}.

\end{itemize}

The publications were not selected only because they used similar terminology. Their task,
input and output, method, dataset, and relevance to the proposed system were considered.
Sources concerned only with general-purpose code generation were excluded when they did not provide insight into command-line interaction or retrieval.

The final set of publications contains both directly related and adjacent work. The
natural-language-to-Bash papers provide the closest comparison with the proposed system.
The retrieval-related papers provide additional context for the decision to identify tools and options from stored descriptions rather than generating unrestricted commands. Together, these sources form the basis of the literature review.

\subsection{Natural-Language-to-Bash Translation}
One of the earliest major datasets for this task was introduced by Lin et al.\ through NL2Bash \cite{lin2018nl2bash}. The objective of NL2Bash is to translate an English description of a task into a corresponding Bash command. The authors created a dataset of natural-language and command pairs and evaluated several neural translation models.
The task is more difficult than ordinary text translation because Bash commands contain several different types of information. A command may include the utility name, multiple option flags, pipelines, filenames, paths, and other user-specific arguments. A model must therefore identify both the general operation and its exact command structure.
NL2Bash distinguishes between the correctness of a complete command and the correctness of its template. A template contains the command structure but replaces specific arguments, such as filenames, with placeholders. The evaluated models achieved better results for templates than for complete commands. This shows that selecting the correct utility and options is already difficult, while handling arbitrary arguments introduces further errors. The authors also identify data sparsity as a major problem. Bash contains many utilities and flags, but individual commands or option combinations may occur only rarely in the training data. A model may therefore fail even when it has seen the utility before. To improve the handling of filenames and other unknown values, NL2Bash uses sub-tokenization and copying mechanisms. These allow parts of the user request to be inserted into the generated command instead of being produced entirely by the model.
NL2Bash demonstrates that natural-language command generation is feasible. However, its best model achieved a top-1 accuracy of approximately 36\% for complete commands and 49\% for command templates \cite{lin2018nl2bash}. This difference highlights the difficulty of generating a fully correct command from unrestricted natural-language input.

The NeurIPS NLC2CMD competition later continued this line of research by providing a shared task for natural-language-to-Bash translation \cite{agarwal2021nlc2cmd}. Competing systems were evaluated on the same hidden test data. The competition encouraged the use of more advanced model architectures and created a common basis for comparing different approaches.

The competition also showed that command translation errors cannot be reduced to one category. A system may select the wrong tool, retrieve the wrong option, omit an argument, or construct the command incorrectly. These error types have different causes and may require different solutions.

\subsection{Improved Command Generation Approaches}
Fu et al.\ proposed NL2CMD as an updated workflow for natural-language-to-Bash translation \cite{fu2023nl2cmd}. The work addresses two limitations of earlier systems: the high effort required to create datasets and the large command vocabulary that translation models must learn.

NL2CMD uses automated command generation and back-translation to create natural-language and Bash pairs. This produces a dataset more than six times larger than earlier datasets while requiring less manual work. The generated commands are executed in an isolated environment to remove commands that fail immediately.

The authors also replace concrete arguments with generic placeholders. For example, filenames and paths are removed before training and inserted again after the command template has been generated. This reduced the Bash vocabulary from 8,184 to 776 tokens and improved the evaluated Transformer models \cite{fu2023nl2cmd}. The model can therefore focus more strongly on command structure, utilities, and options.

The NL2CMD system achieved substantially better results than earlier baselines. However, its output is still generated by a model trained on known command patterns. Support for a new command-line tool would require suitable examples, additional training data, or model adaptation.

Execution-based validation also has limitations. A command that exits successfully is not necessarily semantically correct. It may operate on the wrong file, produce an incomplete result, or perform an unintended action. Execution can therefore remove invalid commands, but it cannot always prove that a command matches the original request. Bharadwaj et al.\ address a related issue in their explainable natural-language-to-Bash system \cite{bharadwaj2021explainable}. Their method represents Bash commands as abstract syntax trees rather than as plain token sequences. This provides the model with explicit information about the structure of utilities, options, and arguments. The system also uses descriptions from manual pages as additional guidance. It aligns parts of the user request with relevant command documentation and produces an explanation together with the generated command. This is intended to make the result easier to understand and to increase trust in the prediction.

The approach achieved results comparable to leading systems from the NLC2CMD competition while also providing explanations \cite{bharadwaj2021explainable}. The work shows that command documentation can support natural-language interpretation. However, the documentation is used to guide a trained generation model. Adding a completely new tool would still require the model to process command structures and documentation in a way that matches its learned representation.

\subsection{Large Language Models and Evaluation Problems}
More recent work has investigated large language models for natural-language-to-Bash translation. Westenfelder et al.\ evaluate LLM-supported translation and identify several problems in existing datasets and evaluation methods \cite{westenfelder2025llm}.
Large language models can generate commands without being trained exclusively on one Bash dataset. They can also interpret more varied language than earlier task-specific systems. However, their outputs remain difficult to evaluate. Two different commands may perform the same operation, while two similar-looking commands may have different effects.
The authors found incorrect examples and unreliable functional-equivalence checks in previous datasets. They therefore corrected and expanded existing data and created a manually verified evaluation set. They also proposed an improved method for estimating whether two Bash commands are functionally equivalent. This work is important because exact text matching is not sufficient for command evaluation. A generated command should not be considered incorrect only because it uses another valid syntax. At the same time, syntactic similarity does not prove that two commands behave identically.
Large language models also introduce practical limitations for local applications. Cloud-based systems require the user request to be sent to an external service. This may be unsuitable when requests contain internal paths, tool names, or other sensitive information. Local models avoid this data transfer, but they can require considerable memory and processing resources.

Another limitation concerns private or niche command-line tools. A general-purpose model can only generate reliable commands for tools represented in its training data or documentation supplied at runtime. Providing new documentation through a retrieval-augmented approach is possible, but the model may still generate options or syntax not present in the retrieved source.

\subsection{Retrieval in Bash-Related Systems}
Most related work generates a complete command directly. BashExplainer differs from these systems because it uses information retrieval as part of its processing pipeline \cite{yu2022bashexplainer}. Its purpose is to generate natural-language comments for existing Bash code rather than to generate commands from user requests.
BashExplainer first retrieves semantically similar Bash code from a repository, and then applies lexical similarity to identify the most relevant example. The retrieved representation is combined with a fine-tuned CodeBERT model and used to generate a description of the command.

The task direction is the reverse of the system developed in this thesis. BashExplainer maps Bash code to natural-language explanations, while this thesis maps natural-language descriptions to command options. Nevertheless, BashExplainer demonstrates that lexical and semantic retrieval can be combined successfully for Bash-related content. Its retrieval process is also relevant because it does not rely only on exact word overlap. Semantic similarity can identify related command behavior, while lexical similarity can preserve exact technical terms. This combination is useful for command-line applications, where both conceptual meaning and specific vocabulary are important.
However, BashExplainer still uses a trained generative model for the final output. It does not retrieve options from a dynamically extensible catalogue, and it is not designed as a small local command-line assistant.

\subsection{Research Gap}
The reviewed systems demonstrate that natural-language requests can be translated into Bash commands. NL2Bash established the task and a corresponding dataset. NLC2CMD provided a shared competition. NL2CMD improved dataset generation and command representation. The AST-based approach added command structure and explanations, while recent LLM-based work improved flexibility and evaluation. These systems primarily treat command assistance as a generation problem. The model receives a natural-language request and produces a complete command: this allows complex and flexible outputs, but it also creates several limitations.

First, generated commands are difficult to validate. A command may be syntactically correct but still fail to perform the intended task. Multiple valid commands may also exist, making exact comparison unreliable.

Second, the systems depend on knowledge learned from training data or supplied through additional context. Supporting an internal or proprietary tool therefore requires new examples, model adaptation, or a retrieval-augmented generation pipeline.

Third, modern generative systems may not be suitable for restricted environments. Cloud-based models require external communication, while large local models can exceed the available memory or acceptable startup time.

Finally, most existing work focuses on generating complete Bash commands for known utilities. Less attention has been given to retrieving individual tools and options from a structured and dynamically extensible source.

This thesis investigates a more restricted approach. Instead of generating unrestricted commands, the system retrieves tools and options from locally stored catalogues. New command-line applications can be added by registering their descriptions, options, keywords, and syntax rules. The matching process combines semantic and lexical retrieval and does not require the base model to be retrained for every new tool.

The restricted output space reduces flexibility but also reduces the possibility of generating unknown utilities or unsupported flags. It is therefore particularly relevant for private, poorly documented, or internally developed command-line tools.
The identified research gap can be summarized as follows: existing research has explored natural-language generation of Bash commands, but it has not sufficiently examined whether a lightweight, local, catalogue-based retrieval system can provide useful command option recommendations for both public and proprietary tools.
The following chapter derives the requirements for such a system and describes its architecture and implementation.
